### 2023-12-09 paper.tex
Initial commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Template for USENIX papers.
%
% History:
%
% - TEMPLATE for Usenix papers, specifically to meet requirements of
%   USENIX '05. originally a template for producing IEEE-format
%   articles using LaTeX. written by Matthew Ward, CS Department,
%   Worcester Polytechnic Institute. adapted by David Beazley for his
%   excellent SWIG paper in Proceedings, Tcl 96. turned into a
%   smartass generic template by De Clarke, with thanks to both the
%   above pioneers. Use at your own risk. Complaints to /dev/null.
%   Make it two column with no page numbering, default is 10 point.
%
% - Munged by Fred Douglis <douglis@research.att.com> 10/97 to
%   separate the .sty file from the LaTeX source template, so that
%   people can more easily include the .sty file into an existing
%   document. Also changed to more closely follow the style guidelines
%   as represented by the Word sample file.
%
% - Note that since 2010, USENIX does not require endnotes. If you
%   want foot of page notes, don't include the endnotes package in the
%   usepackage command, below.
% - This version uses the latex2e styles, not the very ancient 2.09
%   stuff.
%
% - Updated July 2018: Text block size changed from 6.5" to 7"
%
% - Updated Dec 2018 for ATC'19:
%
%   * Revised text to pass HotCRP's auto-formatting check, with
%     hotcrp.settings.submission_form.body_font_size=10pt, and
%     hotcrp.settings.submission_form.line_height=12pt
%
%   * Switched from \endnote-s to \footnote-s to match Usenix's policy.
%
%   * \section* => \begin{abstract} ... \end{abstract}
%
%   * Make template self-contained in terms of bibtex entires, to allow
%     this file to be compiled. (And changing refs style to 'plain'.)
%
%   * Make template self-contained in terms of figures, to
%     allow this file to be compiled. 
%
%   * Added packages for hyperref, embedding fonts, and improving
%     appearance.
%   
%   * Removed outdated text.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix}

% to be able to draw some self-contained figs
\usepackage{tikz}
\usepackage{amsmath}

% inlined bib file
\usepackage{filecontents}

%-------------------------------------------------------------------------------
\begin{filecontents}{\jobname.bib}
%-------------------------------------------------------------------------------
@Book{arpachiDusseau18:osbook,
  author =       {Arpaci-Dusseau, Remzi H. and Arpaci-Dusseau Andrea C.},
  title =        {Operating Systems: Three Easy Pieces},
  publisher =    {Arpaci-Dusseau Books, LLC},
  year =         2015,
  edition =      {1.00},
  note =         {\url{http://pages.cs.wisc.edu/~remzi/OSTEP/}}
}
@InProceedings{waldspurger02,
  author =       {Waldspurger, Carl A.},
  title =        {Memory resource management in {VMware ESX} server},
  booktitle =    {USENIX Symposium on Operating System Design and
                  Implementation (OSDI)},
  year =         2002,
  pages =        {181--194},
  note =         {\url{https://www.usenix.org/legacy/event/osdi02/tech/waldspurger/waldspurger.pdf}}}
\end{filecontents}

% %-------------------------------------------------------------------------------
\begin{document}
% %-------------------------------------------------------------------------------

%don't want date printed
\date{}

% make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Formatting Submissions for a USENIX Conference:\\
  An (Incomplete) Example}

%for single author (just remove % characters)
\author{
{\rm Your N.\ Here}\\
Your Institution
\and
{\rm Second Name}\\
Second Institution
% copy the following lines to add more authors
% \and
% {\rm Name}\\
%Name Institution
} % end author

\maketitle

%-------------------------------------------------------------------------------
\begin{abstract}
%-------------------------------------------------------------------------------
Your abstract text goes here. Just a few facts. Whet our appetites.
Not more than 200 words, if possible, and preferably closer to 150.
\end{abstract}

%-------------------------------------------------------------------------------
\section{Introduction}
%-------------------------------------------------------------------------------

A paragraph of text goes here. Lots of text. Plenty of interesting
text. Text text text text text text text text text text text text text
text text text text text text text text text text text text text text
text text text text text text text text text text text text text text
text text text text text text text.
More fascinating text. Features galore, plethora of promises.

%-------------------------------------------------------------------------------
\section{Footnotes, Verbatim, and Citations}
%-------------------------------------------------------------------------------

Footnotes should be places after punctuation characters, without any
spaces between said characters and footnotes, like so.
\footnote{Remember that USENIX format stopped using endnotes and is
  now using regular footnotes.} And some embedded literal code may
look as follows.

\begin{verbatim}
int main(int argc, char *argv[]) 
{
    return 0;
}
\end{verbatim}

Now we're going to cite somebody. Watch for the cite tag. Here it
comes. Arpachi-Dusseau and Arpachi-Dusseau co-authored an excellent OS
book, which is also really funny~\cite{arpachiDusseau18:osbook}, and
Waldspurger got into the SIGOPS hall-of-fame due to his seminal paper
about resource management in the ESX hypervisor~\cite{waldspurger02}.

The tilde character (\~{}) in the tex source means a non-breaking
space. This way, your reference will always be attached to the word
that preceded it, instead of going to the next line.

And the 'cite' package sorts your citations by their numerical order
of the corresponding references at the end of the paper, ridding you
from the need to notice that, e.g, ``Waldspurger'' appears after
``Arpachi-Dusseau'' when sorting references
alphabetically~\cite{waldspurger02,arpachiDusseau18:osbook}.

It'd be nice and thoughtful of you to include a suitable link in each
and every bibtex entry that you use in your submission, to allow
reviewers (and other readers) to easily get to the cited work, as is
done in all entries found in the References section of this document.

Now we're going take a look at Section~\ref{sec:figs}, but not before
observing that refs to sections and citations and such are colored and
clickable in the PDF because of the packages we've included.

%-------------------------------------------------------------------------------
\section{Floating Figures and Lists}
\label{sec:figs}
%-------------------------------------------------------------------------------

%---------------------------
\begin{figure}
\begin{center}
\begin{tikzpicture}
  \draw[thin,gray!40] (-2,-2) grid (2,2);
  \draw[<->] (-2,0)--(2,0) node[right]{$x$};
  \draw[<->] (0,-2)--(0,2) node[above]{$y$};
  \draw[line width=2pt,blue,-stealth](0,0)--(1,1)
        node[anchor=south west]{$\boldsymbol{u}$};
  \draw[line width=2pt,red,-stealth](0,0)--(-1,-1)
        node[anchor=north east]{$\boldsymbol{-u}$};
\end{tikzpicture}
\end{center}
\caption{\label{fig:vectors} Text size inside figure should be as big as
  caption's text. Text size inside figure should be as big as
  caption's text. Text size inside figure should be as big as
  caption's text. Text size inside figure should be as big as
  caption's text. Text size inside figure should be as big as
  caption's text. }
\end{figure}
%% %---------------------------

Here's a typical reference to a floating figure:
Figure~\ref{fig:vectors}. Floats should usually be placed where latex
wants then. Figure\ref{fig:vectors} is centered, and has a caption
that instructs you to make sure that the size of the text within the
figures that you use is as big as (or bigger than) the size of the
text in the caption of the figures. Please do. Really.

In our case, we've explicitly drawn the figure inlined in latex, to
allow this tex file to cleanly compile. But usually, your figures will
reside in some file.pdf, and you'd include them in your document
with, say, \textbackslash{}includegraphics.

Lists are sometimes quite handy. If you want to itemize things, feel
free:

\begin{description}

\item[fread] a function that reads from a \texttt{stream} into the
  array \texttt{ptr} at most \texttt{nobj} objects of size
  \texttt{size}, returning returns the number of objects read.

\item[Fred] a person's name, e.g., there once was a dude named Fred
  who separated usenix.sty from this file to allow for easy
  inclusion.
\end{description}

\noindent
The noindent at the start of this paragraph in its tex version makes
it clear that it's a continuation of the preceding paragraph, as
opposed to a new paragraph in its own right.

\subsection{LaTeX-ing Your TeX File}
%-----------------------------------

People often use \texttt{pdflatex} these days for creating pdf-s from
tex files via the shell. And \texttt{bibtex}, of course. Works for us.

%-------------------------------------------------------------------------------
\section*{Acknowledgments}
%-------------------------------------------------------------------------------

The USENIX latex style is old and very tired, which is why
there's no \textbackslash{}acks command for you to use when
acknowledging. Sorry.

%-------------------------------------------------------------------------------
\section*{Availability}
%-------------------------------------------------------------------------------

USENIX program committees give extra points to submissions that are
backed by artifacts that are publicly available. If you made your code
or data available, it's worth mentioning this fact in a dedicated
section.

%-------------------------------------------------------------------------------
\bibliographystyle{plain}
\bibliography{\jobname}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%  LocalWords:  endnotes includegraphics fread ptr nobj noindent
%%  LocalWords:  pdflatex acks

### 2023-12-21 paper.tex
Add abstract

@@
 @Book{arpachiDusseau18:osbook,
-  author =       {Arpaci-Dusseau, Remzi H. and Arpaci-Dusseau Andrea C.},
+  author =       {-Dusseau, Remzi H. and Arpaci-Dusseau Andrea C.},
   title =        {Operating Systems: Three Easy Pieces},
@@
 % make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{\Large \bf Formatting Submissions for a USENIX Conference:\\
-  An (Incomplete) Example}
+\title{\Large \bf FlexAlloc: Dynamic Memory Partitioning\\
+    between the Corevisor and Hostvisor in SeKVM}
 
@@
 \author{
-{\rm Your N.\ Here}\\
-Your Institution
+{\rm Matthew Roger Nelson}\\
+Columbia University
 \and
-{\rm Second Name}\\
-Second Institution
-% copy the following lines to add more authors
-% \and
-% {\rm Name}\\
-%Name Institution
+{\rm Ryan Wee}\\
+Columbia University
+\and
+{\rm Zeheng Yang}\\
+Columbia University
 } % end author
@@
 %-------------------------------------------------------------------------------
-Your abstract text goes here. Just a few facts. Whet our appetites.
-Not more than 200 words, if possible, and preferably closer to 150.
+SeKVM is a hypervisor for ARM processors that protects the memory of guest
+virtual machines (VMs) from an untrusted host machine. It comprises a
+\textit{corevisor} that runs at EL2, and a \textit{hostvisor} that runs at EL1
+to leverage the functionality of the untrusted host kernel. However, memory is
+statically partitioned at compile time between the corevisor and the hostvisor
+in SeKVM. This leads to memory wastage when the number of VMs is low, and an
+inability to support the desired number of VMs when the number of VMs is high.
+In this paper, we present FlexAlloc, an extension of SeKVM that allows dynamic
+runtime partitioning of memory between the corevisor and the hostvisor. In
+particular, the memory needed for the stage-2 page tables of a guest VM is
+only allocated to the corevisor when a VM is created. This memory is also
+returned to the hostvisor when a VM is destroyed. Existing implementions
+of \textit{alloc\_pages} do not support allocating a contiguous region of memory
+equal in size to the memory needed for a VM's stage-2 page tables. Hence,
+FlexAlloc also modifies the iterators used to accesss a VM's stage-2 page
+tables, so that the memory used for a VM's stage-2 page tables can be composed
+of noncontiguous regions. Our work ensures that a machine running SeKVM uses
+memory more efficiently, and also allows such machines to support a larger
+number of guest VMs. This allows public cloud providers and organizations
+running their own VMs to scale up their operations at a lower cost.
 \end{abstract}
@@
 
-The USENIX latex style is old and very tired, which is why
-there's no \textbackslash{}acks command for you to use when
-acknowledging. Sorry.
-
-%-------------------------------------------------------------------------------
-\section*{Availability}
-%-------------------------------------------------------------------------------
-
-USENIX program committees give extra points to submissions that are
-backed by artifacts that are publicly available. If you made your code
-or data available, it's worth mentioning this fact in a dedicated
-section.
+Shih-Wei Li, Xupeng Li, Ronghui Gu, Jason Nieh, and John Zhuang Hui
 

### 2023-12-21 paper.tex
Add acknowledgements

@@
 %-------------------------------------------------------------------------------
-@Book{arpachiDusseau18:osbook,
-  author =       {-Dusseau, Remzi H. and Arpaci-Dusseau Andrea C.},
-  title =        {Operating Systems: Three Easy Pieces},
-  publisher =    {Arpaci-Dusseau Books, LLC},
-  year =         2015,
-  edition =      {1.00},
-  note =         {\url{http://pages.cs.wisc.edu/~remzi/OSTEP/}}
+@InProceedings{hypsec,
+  author =       {Li, Shih-Wei and Koh, John S. and Nieh, Jason},
+  title =        {Protecting Cloud Virtual Machines from Commodity Hypervisor and Host Operating System Exploit},
+  booktitle =    {Proceedings of the 28th USENIX Security Symposium},
+  year =         2019,
+  pages =        {1357--1374},
+  note =         {\url{https://www.cs.columbia.edu/~nieh/pubs/security2019_hypsec.pdf}}
 }
-@InProceedings{waldspurger02,
-  author =       {Waldspurger, Carl A.},
-  title =        {Memory resource management in {VMware ESX} server},
-  booktitle =    {USENIX Symposium on Operating System Design and
-                  Implementation (OSDI)},
-  year =         2002,
-  pages =        {181--194},
-  note =         {\url{https://www.usenix.org/legacy/event/osdi02/tech/waldspurger/waldspurger.pdf}}}
+@InProceedings{sekvm,
+  author =       {Li, Shih-Wei and Li, Xupeng and Gu, Ronghui and Nieh, Jason and Hui, John Z.},
+  title =        {Formally Verified Memory Protection for a Commodity Multiprocessor Hypervisor},
+  booktitle =    {Proceedings of the 30th USENIX Security Symposium.},
+  year =         2021,
+  pages =        {3953--3970},
+  note =         {\url{https://www.usenix.org/system/files/sec21-li-shih-wei.pdf}}
+}
+@Unpublished{xuheng,
+    author =     {Li, Xuheng},
+    title =      {Fall 23 Operating Systems 2 Homework},
+    year =       {2023},
+    note =       {\url{https://xuhengli.notion.site}}
+}
 \end{filecontents}
@@
 
-A paragraph of text goes here. Lots of text. Plenty of interesting
-text. Text text text text text text text text text text text text text
-text text text text text text text text text text text text text text
-text text text text text text text text text text text text text text
-text text text text text text text.
-More fascinating text. Features galore, plethora of promises.
+TODO~\cite{sekvm, hypsec}
 
 %-------------------------------------------------------------------------------
-\section{Footnotes, Verbatim, and Citations}
+% \section{Footnotes, Verbatim, and Citations}
 %-------------------------------------------------------------------------------
 
-Footnotes should be places after punctuation characters, without any
-spaces between said characters and footnotes, like so.
-\footnote{Remember that USENIX format stopped using endnotes and is
-  now using regular footnotes.} And some embedded literal code may
-look as follows.
+% Footnotes should be places after punctuation characters, without any
+% spaces between said characters and footnotes, like so.
+% \footnote{Remember that USENIX format stopped using endnotes and is
+%   now using regular footnotes.} And some embedded literal code may
+% look as follows.
 
-\begin{verbatim}
-int main(int argc, char *argv[]) 
-{
-    return 0;
-}
-\end{verbatim}
+% \begin{verbatim}
+% int main(int argc, char *argv[]) 
+% {
+%     return 0;
+% }
+% \end{verbatim}
 
-Now we're going to cite somebody. Watch for the cite tag. Here it
-comes. Arpachi-Dusseau and Arpachi-Dusseau co-authored an excellent OS
-book, which is also really funny~\cite{arpachiDusseau18:osbook}, and
-Waldspurger got into the SIGOPS hall-of-fame due to his seminal paper
-about resource management in the ESX hypervisor~\cite{waldspurger02}.
+% Now we're going to cite somebody. Watch for the cite tag. Here it
+% comes. Arpachi-Dusseau and Arpachi-Dusseau co-authored an excellent OS
+% book, which is also really funny~\cite{arpachiDusseau18:osbook}, and
+% Waldspurger got into the SIGOPS hall-of-fame due to his seminal paper
+% about resource management in the ESX hypervisor~\cite{waldspurger02}.
 
-The tilde character (\~{}) in the tex source means a non-breaking
-space. This way, your reference will always be attached to the word
-that preceded it, instead of going to the next line.
+% The tilde character (\~{}) in the tex source means a non-breaking
+% space. This way, your reference will always be attached to the word
+% that preceded it, instead of going to the next line.
 
-And the 'cite' package sorts your citations by their numerical order
-of the corresponding references at the end of the paper, ridding you
-from the need to notice that, e.g, ``Waldspurger'' appears after
-``Arpachi-Dusseau'' when sorting references
-alphabetically~\cite{waldspurger02,arpachiDusseau18:osbook}.
+% And the 'cite' package sorts your citations by their numerical order
+% of the corresponding references at the end of the paper, ridding you
+% from the need to notice that, e.g, ``Waldspurger'' appears after
+% ``Arpachi-Dusseau'' when sorting references
+% alphabetically~\cite{sekvm,hypsec}.
 
-It'd be nice and thoughtful of you to include a suitable link in each
-and every bibtex entry that you use in your submission, to allow
-reviewers (and other readers) to easily get to the cited work, as is
-done in all entries found in the References section of this document.
+% It'd be nice and thoughtful of you to include a suitable link in each
+% and every bibtex entry that you use in your submission, to allow
+% reviewers (and other readers) to easily get to the cited work, as is
+% done in all entries found in the References section of this document.
 
-Now we're going take a look at Section~\ref{sec:figs}, but not before
-observing that refs to sections and citations and such are colored and
-clickable in the PDF because of the packages we've included.
+% Now we're going take a look at Section~\ref{sec:figs}, but not before
+% observing that refs to sections and citations and such are colored and
+% clickable in the PDF because of the packages we've included.
 
 %-------------------------------------------------------------------------------
-\section{Floating Figures and Lists}
-\label{sec:figs}
+% \section{Floating Figures and Lists}
+% \label{sec:figs}
 %-------------------------------------------------------------------------------
@@
 %---------------------------
-\begin{figure}
-\begin{center}
-\begin{tikzpicture}
-  \draw[thin,gray!40] (-2,-2) grid (2,2);
-  \draw[<->] (-2,0)--(2,0) node[right]{$x$};
-  \draw[<->] (0,-2)--(0,2) node[above]{$y$};
-  \draw[line width=2pt,blue,-stealth](0,0)--(1,1)
-        node[anchor=south west]{$\boldsymbol{u}$};
-  \draw[line width=2pt,red,-stealth](0,0)--(-1,-1)
-        node[anchor=north east]{$\boldsymbol{-u}$};
-\end{tikzpicture}
-\end{center}
-\caption{\label{fig:vectors} Text size inside figure should be as big as
-  caption's text. Text size inside figure should be as big as
-  caption's text. Text size inside figure should be as big as
-  caption's text. Text size inside figure should be as big as
-  caption's text. Text size inside figure should be as big as
-  caption's text. }
-\end{figure}
+% \begin{figure}
+% \begin{center}
+% \begin{tikzpicture}
+%   \draw[thin,gray!40] (-2,-2) grid (2,2);
+%   \draw[<->] (-2,0)--(2,0) node[right]{$x$};
+%   \draw[<->] (0,-2)--(0,2) node[above]{$y$};
+%   \draw[line width=2pt,blue,-stealth](0,0)--(1,1)
+%         node[anchor=south west]{$\boldsymbol{u}$};
+%   \draw[line width=2pt,red,-stealth](0,0)--(-1,-1)
+%         node[anchor=north east]{$\boldsymbol{-u}$};
+% \end{tikzpicture}
+% \end{center}
+% \caption{\label{fig:vectors} Text size inside figure should be as big as
+%   caption's text. Text size inside figure should be as big as
+%   caption's text. Text size inside figure should be as big as
+%   caption's text. Text size inside figure should be as big as
+%   caption's text. Text size inside figure should be as big as
+%   caption's text. }
+% \end{figure}
 %% %---------------------------
 
-Here's a typical reference to a floating figure:
-Figure~\ref{fig:vectors}. Floats should usually be placed where latex
-wants then. Figure\ref{fig:vectors} is centered, and has a caption
-that instructs you to make sure that the size of the text within the
-figures that you use is as big as (or bigger than) the size of the
-text in the caption of the figures. Please do. Really.
+% Here's a typical reference to a floating figure:
+% Figure~\ref{fig:vectors}. Floats should usually be placed where latex
+% wants then. Figure\ref{fig:vectors} is centered, and has a caption
+% that instructs you to make sure that the size of the text within the
+% figures that you use is as big as (or bigger than) the size of the
+% text in the caption of the figures. Please do. Really.
 
-In our case, we've explicitly drawn the figure inlined in latex, to
-allow this tex file to cleanly compile. But usually, your figures will
-reside in some file.pdf, and you'd include them in your document
-with, say, \textbackslash{}includegraphics.
+% In our case, we've explicitly drawn the figure inlined in latex, to
+% allow this tex file to cleanly compile. But usually, your figures will
+% reside in some file.pdf, and you'd include them in your document
+% with, say, \textbackslash{}includegraphics.
 
-Lists are sometimes quite handy. If you want to itemize things, feel
-free:
+% Lists are sometimes quite handy. If you want to itemize things, feel
+% free:
 
-\begin{description}
+% \begin{description}
 
-\item[fread] a function that reads from a \texttt{stream} into the
-  array \texttt{ptr} at most \texttt{nobj} objects of size
-  \texttt{size}, returning returns the number of objects read.
+% \item[fread] a function that reads from a \texttt{stream} into the
+%   array \texttt{ptr} at most \texttt{nobj} objects of size
+%   \texttt{size}, returning returns the number of objects read.
 
-\item[Fred] a person's name, e.g., there once was a dude named Fred
-  who separated usenix.sty from this file to allow for easy
-  inclusion.
-\end{description}
+% \item[Fred] a person's name, e.g., there once was a dude named Fred
+%   who separated usenix.sty from this file to allow for easy
+%   inclusion.
+% \end{description}
 
-\noindent
-The noindent at the start of this paragraph in its tex version makes
-it clear that it's a continuation of the preceding paragraph, as
-opposed to a new paragraph in its own right.
+% \noindent
+% The noindent at the start of this paragraph in its tex version makes
+% it clear that it's a continuation of the preceding paragraph, as
+% opposed to a new paragraph in its own right.
 
-\subsection{LaTeX-ing Your TeX File}
+% \subsection{LaTeX-ing Your TeX File}
 %-----------------------------------
 
-People often use \texttt{pdflatex} these days for creating pdf-s from
-tex files via the shell. And \texttt{bibtex}, of course. Works for us.
+% People often use \texttt{pdflatex} these days for creating pdf-s from
+% tex files via the shell. And \texttt{bibtex}, of course. Works for us.
 
@@
 
-Shih-Wei Li, Xupeng Li, Ronghui Gu, Jason Nieh, and John Zhuang Hui
+Shih-Wei Li, John S. Koh, and Jason Nieh implemented HypSec, which serves as
+the predecessor to SeKVM. Shih-Wei Li, Xupeng Li, Ronghui Gu, Jason Nieh, and
+John Zhuang Hui implemented and formally verfiied SeKVM. Xuheng Li wrote a
+helpful set of notes describing the boot and initialization process in SeKVM,
+as well as the steps needed to build and run SeKVM~\cite{xuheng}. Jason Nieh and
+Xuheng Li provided helpful comments on the project and helped with the design
+on FlexAlloc. This paper was written as part of the class COMS E6118 Operating
+Systems II at Columbia University.
 

### 2023-12-22 paper.tex
Add introduction

@@
 }
-@Unpublished{xuheng,
+@Misc{xuheng,
     author =        {Li, Xuheng},
-    title =      {Fall 23 Operating Systems 2 Homework},
+    title =         {\text{Fall 23 Operating Systems 2 Homework}},
     year =          {2023},
-    note =       {\url{https://xuhengli.notion.site}}
+    howpublished =  {\url{https://xuhengli.notion.site}},
+    note =          {Accessed: 2023-12-21}
+}
+@Misc{arm,
+    author =        {ARM Developer Documentation},
+    title =         {Learn the architecture - AArch64 virtualization},
+    howpublished =  {\url{https://developer.arm.com/documentation/102142/0100/Stage-2-translation}},
+    note =          {Accessed: 2023-12-21}
+}
+@Misc{vmlinux.lds.S,
+    author =        {XuhengLi (GitHub username)},
+    title =         {vmlinux.lds.S in osdi23-paper114-sekvm},
+    howpublished =  {\url{https://github.com/columbia/osdi23-paper114-sekvm/blob/ae/arch/arm64/kernel/vmlinux.lds.S}},
+    note =          {Accessed: 2023-12-21}
+}
+@Misc{kernel-pgtable.h,
+    author =        {XuhengLi (GitHub username)},
+    title =         {kernel-pagetable.h in osdi23-paper114-sekvm},
+    howpublished =  {\url{https://github.com/columbia/osdi23-paper114-sekvm/blob/ae/arch/arm64/include/asm/kernel-pgtable.h}},
+    note =          {Accessed: 2023-12-21}
+}
+@Misc{el1.c,
+    author =        {XuhengLi (GitHub username)},
+    title =         {el1.c in osdi23-paper114-sekvm},
+    howpublished =  {\url{https://github.com/columbia/osdi23-paper114-sekvm/blob/ae/arch/arm64/hypsec_proved/el1.c}},
+    note =          {Accessed: 2023-12-21}       
 }
@@
 virtual machines (VMs) from an untrusted host machine. It comprises a
-\textit{corevisor} that runs at EL2, and a \textit{hostvisor} that runs at EL1
-to leverage the functionality of the untrusted host kernel. However, memory is
-statically partitioned at compile time between the corevisor and the hostvisor
-in SeKVM. This leads to memory wastage when the number of VMs is low, and an
+corevisor that runs at EL2, and a hostvisor that runs at EL1
+to leverage the functionality of the untrusted host kernel. However, SeKVM
+statically partitions memory between the corevisor and hostvisor at compile
+time. This leads to memory wastage when the number of VMs is low, and an
 inability to support the desired number of VMs when the number of VMs is high.
@@
 particular, the memory needed for the stage-2 page tables of a guest VM is
-only allocated to the corevisor when a VM is created. This memory is also
+only allocated to the corevisor when a VM is created. This memory is then
 returned to the hostvisor when a VM is destroyed. Existing implementions
@@
 %-------------------------------------------------------------------------------
-\section{Introduction}
+\section{Introduction and Background}
 %-------------------------------------------------------------------------------
 
-TODO~\cite{sekvm, hypsec}
+Virtual machines (VMs) have become increasingly popular over the past few
+decades. In particular, they improve utilization of existing hardware by
+allowing users to run multiple unmodified operating systems on a single machine.
+A key concern when it comes to VMs is isolation. A single machine may be used to
+run multiple VMs created by different users who generally want their data
+to be protected from one another. As a result, hypervisors generally maintain
+some sort of secure boundary between guest VMs. However, another important
+aspect of isolation is protecting guest VMs from the host machine. A hypervisor
+that fully trusts its host machine is vulnerable to malicious actors who
+gain control of the host operating system itself. This is dangerous because
+the host machine could be used to steal confidential corporate data or sensitive
+personal information.
+
+SeKVM is a formally verified hypervisor for ARM that isolates KVM's trusted computing
+base into a small core~\cite{sekvm}. It builds on the previous work of HypSec,
+which is a hypervisor that protects VMs from the host machine~\cite{hypsec}.
+HypSec consists of two layers. The first layer is an untrusted
+\textit{hostvisor}, which operates at EL1 together with the host operating
+system. The hostvisor leverages many of the data structures and functions
+built into the host operating system for VM management. The second layer is a
+trusted \textit{corevisor}, which operates at the more privileged EL2 level.
+This corevisor is responsible for isolating VMs from each other and from the
+host machine. SeKVM builds on HypSec's design, with the main difference being
+that its trusted computing base is formally verified. In SeKVM, the two layers
+are termed \textit{KServ} and \textit{KCore} respectively. However, in this
+paper, we use the terms hostvisor and corevisor to remain
+consistent with the underlying HypSec kernel source code.
+
+SeKVM protects guest VMs in three main areas: CPU, memory, and I/O. In the
+area of memory, SeKVM leverages ARM hardware support for virtualization. ARM
+allows hypervisors to control memory access using stage-2
+translation~\cite{arm}. In particular, the guest operating systems and host
+operating system each think they are using page tables to map a virtual address space to the
+physical address space. However, in reality, they only control the mapping
+of virtual addresses to intermediate physical addresses. When translating
+virtual addresses, the MMU starts by using these page tables to translate
+virtual addresses into intermediate physical addresses. However, it then 
+uses a second set of page tables to translate
+intermediate phyiscal addresses into the actual machine physical addresses.
+This second set of page tables, called stage-2 page tables, are controlled by
+the hypervisor running at EL2. In the context of SeKVM, the stage-2 page tables
+are managed by the corevisor. When the host operating system tries to
+access some physical memory, the corevisor first checks that this
+physical memory does not belong to a guest VM.
+
+The hostvisor cannot be allowed to access the memory used to store
+the stage-2 page tables of guest VMs. Otherwise, it could simply modify these
+stage-2 page tables so that guest VMs write their data to memory that can be
+accessed and manipulated by the hostvisor. In other words, the memory used for
+these stage-2 page tables must `belong' to the corevisor. In SeKVM,
+the corevisor is statically allocated this memory on bootup. This is done by
+a linker script, \textit{arch/arm64/kernel/vmlinux.lds.S}~\cite{vmlinux.lds.S}, which allocates some
+contiguous region of memory for the stage-2 page tables and demarcates it using the labels \textit{stage2\_pgs\_start}
+and \textit{stage2\_pgs\_end}. The size of this memory region is
+determined by the macro \textit{STAGE2\_PAGES\_SIZE}, which is defined in
+\textit{arch/arm64/include/asm/kernel-pgtable.h}~\cite{kernel-pgtable.h}. In
+particular, this memory region is large enough to store the stage-2 page table
+for the corevisor, the stage-2 page table for the hostvisor,
+and the stage-2 page tables for sixteen guest VMs. The corevisor and hostvisor are
+each allocated sixteen 2M pages for their stage-2 page tables, while each guest VM
+is allocated four 2M pages. Altogether, the corevisor is given 96 2M
+pages for all of these stage-2 page tables. During bootup, the hostvisor
+allocates the beginning of this memory region to the corevisor and itself. It then 
+initializes the corevisor's VM-management data structure such that the remainder of this stage-2 page table memory region
+is evenly divided between each of the sixteen possible guest VMs. In particular, the hostvisor sets
+\textit{el2\_data->vm\_info[i].page\_pool\_start =
+pool\_start + (STAGE2\_VM\_POOL\_SIZE * (i - 1))} for each possible VMID~\cite{el1.c}.
+Here, \textit{vm\_info[i].page\_pool\_start} marks the start of the stage-2
+page table for the \textit{i}th VM.
+
+The key observation here is that this partitioning of memory is done statically
+at compile time. In other words, the hostvisor cannot reclaim the memory it has
+allocated to the corevisor. Similarly, the corevisor cannot ask for more memory
+from the hostvisor. This is in contrast to the dynamic partitioning of memory
+between the hostvisor and guest VMs. Guest VMs can request memory from the
+hostvisor via a page fault, and the hostvisor can reclaim memory from guest VMs
+using ballooning~\cite{hypsec}. The static partitioning of memory between the
+hostvisor and the corevisor is undesirable for two reasons:
+
+\begin{itemize}
+    \item First, unused memory cannot be reclaimed from the corevisor. Say
+    the number of VMs being run is $n$, such that $n < 16$. Then there will be
+    $(16 - n) * 4$ 2M pages reserved for the stage-2 page tables of guest VMs that
+    do not exist. Dynamic partitioning of memory would allow this memory to be reclaimed by the hostvisor, to
+    alleviate memory pressure in the host. This memory could also be given to guest VMs, to
+    alleviate memory pressure in these VMs. Giving VMs more memory would improve
+    VM performance, because less data would need to be moved to the swap partition. 
+    \item Second, there is a hard upper limit on the number of VMs that can
+    be supported by SeKVM. Users cannot add VMs above the limit of sixteen,
+    because there will be no memory in the corevisor to support the stage-2
+    page tables of these VMs. Users could always choose to statically allocate
+    even more memory to the corevisor upon bootup, but this would exacerbate
+    the problem of unused memory mentioned above.
+\end{itemize}
+
+Static allocation of memory to the corevisor is not limited to memory for the
+stage-2 page tables of each VM. In addition to this, SeKVM also statically
+allocates memory to the corevisor to store the metadata of each VM. The same linker script
+mentioned above reserves another contiguous region of memory equal in size to 32
+2M pages, and demarcates it using \textit{el2\_data\_start}
+and \textit{el2\_data\_end}~\cite{vmlinux.lds.S}. This memory is given to the
+corevisor on bootup. Part of this memory is used to
+store an array of \textit{struct el2\_vm\_info} objects, where each object is
+used to store the metadata of each guest VM. SeKVM could potentially also benefit
+from dynamic allocation of the memory used to store this VM metadata, where
+this memory is only given to the corevisor upon VM creation and returned to the
+hostvisor upon VM deletion. However,
+we chose not to focus on this because the size of each \textit{struct el2\_vm\_info} object
+is 768 bytes. The total size of all of the \textit{struct el2\_vm\_info} objects
+is 12288 bytes, which is less than four 4K pages in total. As a result, the
+problem of unused memory is not as significant, since we waste at most four
+4K pages of memory. In any case, memory can only be allocated at the
+page granularity. Hence, in the case where we want to create a new VM and there is no 
+more space in the existing dynamically-allocated memory for the new \textit{struct el2\_vm\_info} object, we would
+inevitably end up allocating an entirely new 4K page, the majority of which would remain unused. (This
+is not the case with VM stage-2 page tables, since the size of the stage-2 page
+table for each VM is nicely page-aligned.) The problem of a hard upper limit
+is also not as significant, since we could easily make the array larger without
+incurring much memory overhead.
+
+%-------------------------------------------------------------------------------
+\section{Design}
+%-------------------------------------------------------------------------------
+
+TODO
+
+%-------------------------------------------------------------------------------
+\section{Implementation}
+%-------------------------------------------------------------------------------
+
+TODO
+
+%-------------------------------------------------------------------------------
+\section{Evaluation}
+%-------------------------------------------------------------------------------
+
+TODO
+
+%-------------------------------------------------------------------------------
+\section{Future Work}
+%-------------------------------------------------------------------------------
+
+TODO
+
+%-------------------------------------------------------------------------------
+\section{Conclusions}
+%-------------------------------------------------------------------------------
+
+TODO
 

### 2023-12-22 paper.tex
Add design and implementation, sans s2 page table iterators

@@
     howpublished =  {\url{https://github.com/columbia/osdi23-paper114-sekvm/blob/ae/arch/arm64/hypsec_proved/el1.c}},
+    note =          {Accessed: 2023-12-21}       
+}
+@Misc{kerneldocs,
+    author =        {kernel.org},
+    title =         {Chapter 6: Physical Page Allocation},
+    howpublished =  {\url{https://www.kernel.org/doc/gorman/html/understand/understand009.html}},
+    note =          {Accessed: 2023-12-22}
+}
+@Misc{page_alloc.c,
+    author =        {Elixir Bootlin},
+    title =         {page\_alloc.c in Linux v5.4.55},
+    howpublished =  {\url{https://elixir.bootlin.com/linux/v5.4.55/source/mm/page_alloc.c#L4718}},
+    note =          {Accessed: 2023-12-22}
+}
+@Misc{mmzone.h,
+    author =        {Elixir Bootlin},
+    title =         {mmzone.h in Linux v5.4.55},
+    howpublished =  {\url{https://elixir.bootlin.com/linux/v5.4.55/source/include/linux/mmzone.h#L27}},
+    note =          {Accessed: 2023-12-22}
+}
+@Misc{el2.c,
+    author =        {XuhengLi (GitHub username)},
+    title =         {el2.c in osdi23-paper114-sekvm},
+    howpublished =  {\url{https://github.com/columbia/osdi23-paper114-sekvm/blob/ae/arch/arm64/hypsec_proved/el2.c}},
     note =          {Accessed: 2023-12-21}       
@@
 4K pages of memory. In any case, memory can only be allocated at the
-page granularity. Hence, in the case where we want to create a new VM and there is no 
-more space in the existing dynamically-allocated memory for the new \textit{struct el2\_vm\_info} object, we would
-inevitably end up allocating an entirely new 4K page, the majority of which would remain unused. (This
+page granularity. Hence, in the case where we want to create a new VM and the new \textit{struct el2\_vm\_info} object
+cannot be placed into the pages that have already been dynamically allocated to the corevisor, we will
+inevitably end up having to allocate an entirely new 4K page. The majority of this new 4K page would then remain unused, until another VM is created. This
 is not the case with VM stage-2 page tables, since the size of the stage-2 page
-table for each VM is nicely page-aligned.) The problem of a hard upper limit
+table for each VM is nicely page-aligned. With the \textit{struct el2\_vm\_info} objects, the problem of a hard upper limit
 is also not as significant, since we could easily make the array larger without
@@
 %-------------------------------------------------------------------------------
-\section{Design}
+\section{Design and Implementation}
 %-------------------------------------------------------------------------------
 
-TODO
+All of the source code described below can be found at \url{https://github.com/nelsonm2991/host-sekvm-project}.
 
-%-------------------------------------------------------------------------------
-\section{Implementation}
-%-------------------------------------------------------------------------------
+\subsection{Modifying the page table iterators}
 
-TODO
+As described above, the original implementation of SeKVM statically
+allocates sixteen contiguous 8M regions to the corevisor. When a VM is created,
+the corevisor uses one of these contiguous 8M regions for the VM's stage-2
+page table. To implement a system of dynamic allocation, we want the hostvisor
+to dynamically reserve a contiguous 8M region at runtime when a VM is created, and then
+pass this memory region to the corevisor. The kernel macro \textit{alloc\_pages}
+is responsible for allocating a large number of contiguous physical pages~\cite{kerneldocs}.
+Examining the implementation of this macro reveals that it eventually
+calls \textit{\_\_alloc\_pages\_nodemask}, which checks that the order requested
+is strictly less than \textit{MAX\_ORDER}~\cite{page_alloc.c}. This
+is in turn defined to be 11~\cite{mmzone.h}. This means that we cannot allocate
+a contiguous region of memory larger than $2^{10} * 4K = 4M$.
+
+To get around this limitation, there are two possible approaches. The first approach
+is to halve the size of the VM's stage-2 page table, so that it can fit within
+a single contiguous 4M region of memory. Alternatively, we can fragment the VM's stage-2
+page table such that it can be composed of non-contiguous regions of memory. We
+chose to implement the latter approach to avoid degrading VM performance. In
+particular, we fragmented the VM's stage-2 page table so that it could be composed
+of eight distinct 1M regions of memory. Each of these regions must be contiguous,
+but each region is distinct and need not be allocated contiguously with the
+other 1M regions. As a result, the hostvisor can easily use \textit{alloc\_pages}
+to allocate the memory needed for a guest VM's stage-2 page tables.
+
+We implement this fragmentation by manipulating the iterators used to traverse
+the guest VM's stage-2 page table. In particular, the VM's stage-2 page table
+is a four-level page table, comprising a page global directory (PGD), a
+page upper directory (PUD), a page middle directory (PMD), and multiple page table
+entries (PTEs). TODO
+
+\subsection{Allocating memory to the corevisor}
+
+FlexAlloc dynamically allocates eight 1M memory regions to the corevisor when a
+VM is created. It does so by passing the base addresses of these eight 1M
+regions as arguments to the \textit{HVC\_REGISTER\_KVM} hypercall. An alternative
+implementation that we considered was creating a new \textit{HVC\_PASS\_MEM} hypercall to pass these
+1M regions to the corevisor, and getting the hostvisor to call \textit{HVC\_PASS\_MEM}
+before calling \textit{HVC\_REGISTER\_KVM}. However, we ultimately chose to
+discard this alternative design because it leads to a more complicated
+implementation. In particular, the corevisor would have to verify a one-to-one
+correspondence between calls to \textit{HVC\_PASS\_MEM} and \textit{HVC\_REGISTER\_KVM}.
+It would also have to handle malicious hostvisors that try to call 
+\textit{HVC\_REGISTER\_KVM} without calling \textit{HVC\_PASS\_MEM}.
+
+The \textit{HVC\_REGISTER\_KVM} hypercall eventually calls \textit{register\_kvm}
+~\cite{el2.c}. Note that at this point we are in the corevisor, running at EL2.
+Within this function, we preprocess the eight 1M regions to ensure that the
+calls to \textit{alloc\_pages} in the hostvisor were successful, and to verify
+that the memory regions provided actually belong to the hostvisor. We then set
+the ownership of these memory regions to the corevisor, so that the host can
+no longer access or manipulate the guest VM's stage-2 page tables. In particular,
+we do the following:
+
+\begin{verbatim}
+for (i = 0; i < 8; ++i) {
+    addr = page_starts[i];
+    end = page_starts[i] + SZ_1M;
+
+    /* ... */
+
+    if (addr == 0) {
+        // Hostvisor's alloc_pages() failed.
+        __hyp_panic();
+    }
+
+    // Change the ownership of these pages
+    // to the corevisor.
+    do {
+        index = get_s2_page_index(addr);
+        owner = get_s2_page_vmid(index);
+
+        if (owner != HOSTVISOR) {
+            // The hostvisor shouldn't be
+            // giving memory that belongs
+            // to the corevisor or to a VM.
+            __hyp_panic();
+        }
+
+        set_s2_page_vmid(index, COREVISOR);
+        addr += PAGE_SIZE;
+        ++page_cnt;
+    } while (addr < end);
+}
+\end{verbatim}
+
+We then build the VM's stage-2 page table using these eight 1M regions, by
+populating the relevant fields in the \textit{struct el2\_data} object used
+by the corevisor for VM management.
+
+\subsection{Reclaiming memory from the corevisor}
+
+When a VM is destroyed, FlexAlloc reclaims the eight 1M memory regions that were
+previously allocated to the corevisor for the VM's stage-2 page table. Returning
+this memory to the hostvisor ensures that it does not remain unused in corevisor
+memory. Unlike the case of VM creation, there is no hypercall made directly to
+the corevisor when a VM is destroyed. Hence, we create a new \textit{HVC\_DESTROY\_KVM}
+hypercall that mirrors the \textit{HVC\_REGISTER\_KVM} hypercall used during VM
+creation. In particular, on ARM architectures, if the \textit{CONFIG\_VERIFIED\_KVM} macro is defined,
+we call \textit{hypsec\_arch\_destroy\_vm}. This function in turn calls \textit{hypsec\_destroy\_kvm},
+which ultimately invokes \textit{HVC\_REGISTER\_KVM} hypercall with the corresponding
+VMID. We do not pass the base addresses of the eight 1M memory regions as arguments
+to this hypercall, since the corevisor can access them directly by reading the
+relevant fields of the \textit{struct el2\_data} object. Once we are in the
+corevisor running in EL2, we return the eight memory regions to the hostvisor.
+We then return to the hostvisor running in EL1, which can then call \textit{\_\_free\_pages}
+on each of the eight 1M regions.
 
@@
 
-TODO
+Not verified
+
+Doesn't memset, doesn't verify the eight 1M regions don't overlap
+
+One key shortcoming of our project is that it does not fully implement
 
@@
 
-TODO
+We have designed and implemented FlexAlloc, a system for
 
@@
 
-% \subsection{LaTeX-ing Your TeX File}
-%-----------------------------------
-
-% People often use \texttt{pdflatex} these days for creating pdf-s from
-% tex files via the shell. And \texttt{bibtex}, of course. Works for us.
-
 %-------------------------------------------------------------------------------
@@
 the predecessor to SeKVM. Shih-Wei Li, Xupeng Li, Ronghui Gu, Jason Nieh, and
-John Zhuang Hui implemented and formally verfiied SeKVM. Xuheng Li wrote a
+John Zhuang Hui implemented and formally verified SeKVM. Xuheng Li wrote a
 helpful set of notes describing the boot and initialization process in SeKVM,
@@
 Xuheng Li provided helpful comments on the project and helped with the design
-on FlexAlloc. This paper was written as part of the class COMS E6118 Operating
+of FlexAlloc. This paper was written as part of the class COMS E6118 Operating
 Systems II at Columbia University.

### 2023-12-22 paper.tex
Add shortcomings and future work, sans DPM

@@
     howpublished =  {\url{https://github.com/columbia/osdi23-paper114-sekvm/blob/ae/arch/arm64/hypsec_proved/el2.c}},
-    note =          {Accessed: 2023-12-21}       
+    note =          {Accessed: 2023-12-22}       
+}
+@Misc{hypsec_host.h,
+    author =        {XuhengLi (GitHub username)},
+    title =         {hypsec\_host.h in osdi23-paper114-sekvm},
+    howpublished =  {\url{https://github.com/columbia/osdi23-paper114-sekvm/blob/ae/arch/arm64/include/asm/hypsec_host.h}},
+    note =          {Accessed: 2023-12-22}       
 }
@@
 the corevisor is statically allocated this memory on bootup. This is done by
-a linker script, \textit{arch/arm64/kernel/vmlinux.lds.S}~\cite{vmlinux.lds.S}, which allocates some
+a linker script, which allocates some
 contiguous region of memory for the stage-2 page tables and demarcates it using the labels \textit{stage2\_pgs\_start}
-and \textit{stage2\_pgs\_end}. The size of this memory region is
-determined by the macro \textit{STAGE2\_PAGES\_SIZE}, which is defined in
-\textit{arch/arm64/include/asm/kernel-pgtable.h}~\cite{kernel-pgtable.h}. In
+and \textit{stage2\_pgs\_end}~\cite{vmlinux.lds.S}. The size of this memory region is
+determined by the macro \textit{STAGE2\_PAGES\_SIZE}~\cite{kernel-pgtable.h}. In
 particular, this memory region is large enough to store the stage-2 page table
@@
 hostvisor upon VM deletion. However,
-we chose not to focus on this because the size of each \textit{struct el2\_vm\_info} object
+we choose not to focus on this because the size of each \textit{struct el2\_vm\_info} object
 is 768 bytes. The total size of all of the \textit{struct el2\_vm\_info} objects
@@
 calls \textit{\_\_alloc\_pages\_nodemask}, which checks that the order requested
-is strictly less than \textit{MAX\_ORDER}~\cite{page_alloc.c}. This
+is strictly less than \textit{MAX\_ORDER}~\cite{page_alloc.c}. \textit{MAX\_ORDER}
 is in turn defined to be 11~\cite{mmzone.h}. This means that we cannot allocate
-a contiguous region of memory larger than $2^{10} * 4K = 4M$.
+a contiguous region of memory larger than $2^{10} * 4K = 4M$. This
+limit is set by the kernel because \textit{alloc\_pages} uses the buddy allocator,
+which recursively splits available memory into half to find the best possible position for
+a contiguous region of memory~\cite{kerneldocs}. However, this also makes it
+undesirable to continuously allocate huge contiguous regions of memory.
 
@@
 a single contiguous 4M region of memory. Alternatively, we can fragment the VM's stage-2
-page table such that it can be composed of non-contiguous regions of memory. We
-chose to implement the latter approach to avoid degrading VM performance. In
-particular, we fragmented the VM's stage-2 page table so that it could be composed
-of eight distinct 1M regions of memory. Each of these regions must be contiguous,
+page table such that it can be composed of multiple non-contiguous regions of memory, whose sizes in total add up to 8M. We
+choose to implement the latter approach to avoid degrading VM performance, and to avoid
+breaking the current implementation of the buddy allocator. In
+particular, we fragment the VM's stage-2 page table so that it can be composed
+of eight distinct 1M regions of memory. Each of these 1M regions must be contiguous,
 but each region is distinct and need not be allocated contiguously with the
@@
 page upper directory (PUD), a page middle directory (PMD), and multiple page table
-entries (PTEs). TODO
+entries (PTEs). TODO (Matthew - Page table fragmentation)
 
@@
 1M regions to the corevisor, and getting the hostvisor to call \textit{HVC\_PASS\_MEM}
-before calling \textit{HVC\_REGISTER\_KVM}. However, we ultimately chose to
+before calling \textit{HVC\_REGISTER\_KVM}. However, we ultimately choose to
 discard this alternative design because it leads to a more complicated
@@
 It would also have to handle malicious hostvisors that try to call 
-\textit{HVC\_REGISTER\_KVM} without calling \textit{HVC\_PASS\_MEM}.
+\textit{HVC\_REGISTER\_KVM} without calling \textit{HVC\_PASS\_MEM}. Morevoer,
+adding an extra hypercall increases latency because it leads to more frequent
+world-switches between the hostvisor running in EL1 and the corevisor running in EL2.
 
@@
 
-We then build the VM's stage-2 page table using these eight 1M regions, by
+\noindent We then build the VM's stage-2 page table using these eight 1M regions, by
 populating the relevant fields in the \textit{struct el2\_data} object used
@@
 we call \textit{hypsec\_arch\_destroy\_vm}. This function in turn calls \textit{hypsec\_destroy\_kvm},
-which ultimately invokes \textit{HVC\_REGISTER\_KVM} hypercall with the corresponding
+which ultimately invokes the \textit{HVC\_REGISTER\_KVM} hypercall with the corresponding
 VMID. We do not pass the base addresses of the eight 1M memory regions as arguments
@@
 
-TODO
+We evaluate FlexAlloc by TODO
 
 %-------------------------------------------------------------------------------
-\section{Future Work}
+\section{Shortcomings and Future Work}
 %-------------------------------------------------------------------------------
 
-Not verified
+One key shortcoming of FlexAlloc is that it is not formally verified. In
+this sense, FlexAlloc is less like SeKVM and more like HypSec. It provides isolation guarantees
+in principle, but cannot be logically proven to be secure. In particular, there
+are a few major flaws in FlexAlloc when it comes to isolating guest VMs from
+the host machine. For instance, the current implementation of FlexAlloc
+does not use \textit{memset} to zero out the memory given to it by the hostvisor
+for the VM's stage-2 page table. This presents a danger if the MMU tries to
+read from a page table entry that should not actually exist, but that has
+previously been written into the memory region by the hostvisor. Another flaw is that
+FlexAlloc does not verify that the eight 1M regions are actually distinct and
+do not overlap with one another. Hence, FlexAlloc still fundamentally relies on
+a non-malicious hostvisor, and would require further patches to be fully secure.
 
-Doesn't memset, doesn't verify the eight 1M regions don't overlap
+Another key shortcoming of our project is that it does not fully implement
+dynamic memory partitioning between the corevisor and hostvisor. As mentioned in
+the introduction, another area in which corevisor memory scales with the number
+of VMs is in maintaining \textit{struct el2\_vm\_info} objects that record the
+metadata associated with each guest VM. Currently, we still statically allocate
+a fixed-size array of \textit{struct el2\_vm\_info} objects in corevisor
+memory ~\cite{hypsec_host.h}. This essentially preserves the hard cap on the
+number of guest VMs that can be run. An easy fix would be to keep this array
+as a statically-allocated array, while increasing the number of elements. Since each
+\textit{struct el2\_vm\_info} object is only 768 bytes, this would not waste a 
+significant amount of memory even if fewer than sixteen VMs are created.
 
-One key shortcoming of our project is that it does not fully implement
+We also do not fully evaluate our implementation by trying to run a large number
+of VMs. To do this, we would have to modify the macros used to define the current
+sixteen-VM limit. It is entirely conceivable that the hostvisor might not be able
+to allocate 1M contiguous regions of memory past a certain point if memory
+pressure and external fragmentation are high. In this case, the hostvisor would
+not be able to allocate the memory regions needed for the corevisor to create
+the stage-2 page table for a new VM. We could perhaps delay this limit by breaking
+the VM stage-2 page tables down into even smaller regions of memory, but this would
+likely lead to more complicated logic when traversing the page table.
+
+Another possible area of future work is exploring other allocators designed for
+allocating large chunks of contiguous memory. We fragment the VM stage-2 page tables
+into eight distinct 1M pages, because \textit{alloc\_pages} does not allow us
+to allocate contiguous regions larger than 4M. It might be possile to avoid this
+by looking at other Linux memory allocators explicitly designed for the purpose
+of large contiguous allocations.
+
+TODO (Matthew - DPM)
 
@@
 
-We have designed and implemented FlexAlloc, a system for
+We have designed and implemented FlexAlloc, a system for dynamic memory
+partitioning between the corevisor and hostvisor in SeKVM.
 
 %-------------------------------------------------------------------------------
-% \section{Footnotes, Verbatim, and Citations}
+\section*{Acknowledgments}
 %-------------------------------------------------------------------------------
 
-% Footnotes should be places after punctuation characters, without any
-% spaces between said characters and footnotes, like so.
-% \footnote{Remember that USENIX format stopped using endnotes and is
-%   now using regular footnotes.} And some embedded literal code may
-% look as follows.
+Shih-Wei Li, John S. Koh, and Jason Nieh implemented HypSec, which serves as
+the predecessor to SeKVM. Shih-Wei Li, Xupeng Li, Ronghui Gu, Jason Nieh, and
+John Zhuang Hui implemented and formally verified SeKVM. Xuheng Li wrote a
+helpful set of notes describing the boot and initialization process in SeKVM,
+as well as the steps needed to build and run SeKVM~\cite{xuheng}. Jason Nieh and
+Xuheng Li provided helpful comments on the project and helped with the design
+of FlexAlloc. This paper was written as part of the class COMS E6118 Operating
+Systems II at Columbia University.
 
-% \begin{verbatim}
-% int main(int argc, char *argv[]) 
-% {
-%     return 0;
-% }
-% \end{verbatim}
+%-------------------------------------------------------------------------------
+\bibliographystyle{plain}
+\bibliography{\jobname}
 
-% Now we're going to cite somebody. Watch for the cite tag. Here it
-% comes. Arpachi-Dusseau and Arpachi-Dusseau co-authored an excellent OS
-% book, which is also really funny~\cite{arpachiDusseau18:osbook}, and
-% Waldspurger got into the SIGOPS hall-of-fame due to his seminal paper
-% about resource management in the ESX hypervisor~\cite{waldspurger02}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\end{document}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% The tilde character (\~{}) in the tex source means a non-breaking
-% space. This way, your reference will always be attached to the word
-% that preceded it, instead of going to the next line.
-
-% And the 'cite' package sorts your citations by their numerical order
-% of the corresponding references at the end of the paper, ridding you
-% from the need to notice that, e.g, ``Waldspurger'' appears after
-% ``Arpachi-Dusseau'' when sorting references
-% alphabetically~\cite{sekvm,hypsec}.
-
-% It'd be nice and thoughtful of you to include a suitable link in each
-% and every bibtex entry that you use in your submission, to allow
-% reviewers (and other readers) to easily get to the cited work, as is
-% done in all entries found in the References section of this document.
-
-% Now we're going take a look at Section~\ref{sec:figs}, but not before
-% observing that refs to sections and citations and such are colored and
-% clickable in the PDF because of the packages we've included.
+%%  LocalWords:  endnotes includegraphics fread ptr nobj noindent
+%%  LocalWords:  pdflatex acks
 
@@
 % \end{description}
-
-% \noindent
-% The noindent at the start of this paragraph in its tex version makes
-% it clear that it's a continuation of the preceding paragraph, as
-% opposed to a new paragraph in its own right.
-
-%-------------------------------------------------------------------------------
-\section*{Acknowledgments}
-%-------------------------------------------------------------------------------
-
-Shih-Wei Li, John S. Koh, and Jason Nieh implemented HypSec, which serves as
-the predecessor to SeKVM. Shih-Wei Li, Xupeng Li, Ronghui Gu, Jason Nieh, and
-John Zhuang Hui implemented and formally verified SeKVM. Xuheng Li wrote a
-helpful set of notes describing the boot and initialization process in SeKVM,
-as well as the steps needed to build and run SeKVM~\cite{xuheng}. Jason Nieh and
-Xuheng Li provided helpful comments on the project and helped with the design
-of FlexAlloc. This paper was written as part of the class COMS E6118 Operating
-Systems II at Columbia University.
-
-%-------------------------------------------------------------------------------
-\bibliographystyle{plain}
-\bibliography{\jobname}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\end{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%  LocalWords:  endnotes includegraphics fread ptr nobj noindent
-%%  LocalWords:  pdflatex acks

### 2023-12-22 paper.tex
Add conclusion

@@
 We have designed and implemented FlexAlloc, a system for dynamic memory
-partitioning between the corevisor and hostvisor in SeKVM.
+partitioning between the corevisor and hostvisor in SeKVM. In particular, we
+modify SeKVM so that the hostvisor allocates memory to the corevisor when a VM is
+created, and reclaims memory from the corevisor when a VM is destroyed. As a
+result, the memory for a VM's stage-2 page table does not need to be statically
+allocated to the corevisor upon bootup. This system reduces the wastage of memory
+when VM demand is low, and increases the number of VMs that can be run
+on a single machine when VM demand is high. We also make a distinct contribution in terms of fragmenting
+a VM's stage-2 page table so that it can be composed of multiple noncontiguous
+regions of memory, instead of a single contiguous 8M region. Altogether, this
+system improves hardware utilization and reduces VM operating costs.
 

### 2023-12-22 paper.tex
Add barebones evaluation

@@
 All of the source code described below can be found at \url{https://github.com/nelsonm2991/host-sekvm-project}.
+Our implementation of FlexAlloc is based on SeKVM, which is in turn based on Linux v5.4.55.
 
@@
 
-We evaluate FlexAlloc by TODO
+We evaluate FlexAlloc by running it on a QEMU-emulated Raspberry Pi 4. In
+particular, we run VMWare Fusion Professional Version 13.0.2 on an M1 Macbook
+Air running MacOS Ventura 13.6. Using VMWare, we create a VM
+with 2 processor cores and 8704 MB memory running Ubuntu 22.04.3 LTS. On this
+VM, we run QEMU 6.2.0 to emulate a Raspberry Pi 4. This emulated Raspberry
+Pi 4 serves as our host machine, on which we run FlexAlloc. Our guest VMs 
+run vanilla Linux v5.4.
+
+Using FlexAlloc, we boot up a guest VM and verify that it is able to start up
+without any kernel panics. We use \textit{print\_string} to verify that we
+are passing eight distinct 1M regions to the corevisor when the VM is created.
+We test a case where the 1M regions are not contiguous, and verify that our
+modified page table iterators are able to successfully handle a VM stage-2 page
+table that is non-contiguous. We also shut down the guest VM and verify that
+all eight 1M regions are returned to the hostvisor. We compare the base address
+of each region allocated during VM creation against the base address of each
+region reclaimed during VM shutdown, and verify that they match each other.
 
@@
 %-------------------------------------------------------------------------------
+
+\subsection{Shortcomings in design}
 
@@
 
-Another key shortcoming of our project is that it does not fully implement
+Another key shortcoming of our design is that it does not fully implement
 dynamic memory partitioning between the corevisor and hostvisor. As mentioned in
@@
 
-We also do not fully evaluate our implementation by trying to run a large number
+\subsection{Shortcomings in evaluation}
+
+We also do not fully evaluate our implementation. For one, we do not stress-test our implementation by running a large number
 of VMs. To do this, we would have to modify the macros used to define the current
@@
 the VM stage-2 page tables down into even smaller regions of memory, but this would
-likely lead to more complicated logic when traversing the page table.
+likely lead to more complicated logic when traversing the page table. In addition, we do not verify whether fragmenting the stage-2 page table results
+in any degradation in VM performance. Ideally, we would run a memory-intensive
+benchmark on our guest VMs to verify that our iterators work properly under a
+high memory load. Finally, we do not quantify the change in corevisor memory usage.
+We assume that by not allocating the memory for VM stage-2 page tables at bootup,
+corevisor memory will automatically be reduced by an amount matching the
+size of this static allocation. This should be verified using actual data. For
+instance, we could periodically iterate over every page and see how many pages
+are owned by the corevisor before and after VM creation.
 
-Another possible area of future work is exploring other allocators designed for
+\subsection{Future work}
+
+A possible area of future work is exploring other allocators designed for
 allocating large chunks of contiguous memory. We fragment the VM stage-2 page tables

### 2023-12-23 paper.tex
Filled in TODOs

@@
 
-We implement this fragmentation by manipulating the iterators used to traverse
-the guest VM's stage-2 page table. In particular, the VM's stage-2 page table
-is a four-level page table, comprising a page global directory (PGD), a
-page upper directory (PUD), a page middle directory (PMD), and multiple page table
-entries (PTEs). TODO (Matthew - Page table fragmentation)
+We implement this fragmentation by manipulating the iterators that provide the next useable
+page at a given level when traversing the stage-2 page table for the purpose of recording
+an additional entry. In the HypSec implementation, each stage-2 page level's pages are
+kept contiguous to one another and never intermixed with pages from other levels.
+Each iterator starts at a particular offset into the VM's statically allocated stage-2
+page table region, and increments each time a new page is consumed during a traversal
+for use at that level. The design of these iterators means that they are limited to
+only operating within a large, contiguous region.
+
+The HypSec page global directory (PGD), page upper directory (PUD), and page middle directory
+(PMD) iterators operate within a single contiguous 8MB region divided into varying sized
+sub-regions for each level. The PGD iterator operates over a 16 * PAGE_SIZE region, the PUD
+iterator over a SZ_2M - (16 * PAGE_SIZE) region, and the PMD iterator over a 3 * SZ_2M region.
+Our implementation eliminates the need for all of these subregions to be contiguous to
+one another and enables fragmentation within each subregion itself.
+
+The FlexAlloc iterators, which can be thought of as page allocators specific to VM stage-2
+page tables, operate over a pool of eight, non-contiguous 1MB regions. The FlexAlloc iterators
+first exhaust all pages from a given 1MB region. Once the current region is exhausted, FlexAlloc
+will jump the iterators to the next region in the pool. Subsequent traversals of the
+stage-2 page tables with the alloc flag enabled will pull pages from the seperate,
+non-contiguous region. By removing the requirement for contiguous 8MB regions, FlexAlloc
+enables reliable dynamic allocation of stage-2 page table regions.
+
+We implemented two versions of the FlexAlloc iterators. In the first version, we used a single,
+generalized iterator shared across all levels of the stage-2 page tables. In the HypSec implementation,
+the iterators are structured such that a PUD page will never be sandwiches between two PMD pages.
+The first version of the FlexAlloc iterators ignores this semantic and instead is non-discrimiant
+in regards to the particular level a page will be used in. We decided to stop development of this
+version before it was fully functional in favor of version two which more closely resembles recent changes made to SeKVM's
+stage-2 page table allocation scheme.
+
+In the second version, we opt to enforce the semantic that pages at each specific level are
+allocated from seperate pools and thus never mixed with one another. To accomplish this while still maintaining the same pool of non-contiguous 1MB regions,
+we split the single, general iterator into nine distinct iterators. One iterator is used
+for PUD pages, two iterators are used for PMD pages, and six iterators are used for PTE pages.
+The number of iterators assigned to each level is based on the potential maximum number of
+pages that could be consumed at that level along with how many distinct 1MB regions each level
+can be allocated from. When a page is allocated at a particular level, the correct iterator is chosen based on
+it's assigned level and whether or not its preceeding iterators have been exhausted.
+
+The naming semantics of the iterators at each level can be confusing so we feel it is worth
+clarifying here. The PGD iterator allocates pages to use as entries inside of a PGD page. This
+implies that the PGD iterator, as an example, actually allocates pages that are used in the
+subsequent level, which in this case is PUD. All iterators follow the pattern of allocating
+pages for a given level to point to which are implicitly actually pages in the next level
+of the stage-2 page tables.
 
@@
 
-TODO (Matthew - DPM)
+To enable dynamic allocation of EL2 memory in general and beyond just the scope of just
+the stage-2 page tables, we proposed and designed, but did not complete the implementation for,
+a dynamic page manager (DPM) interface. DPM would enable other "services" within EL2 to
+utilize dynamic allocation for their memory regions without needing to handle the process
+of receiving memory from the host.
 

### 2023-12-23 paper.tex
Some formatting fixes

--- 2023-12-22 paper.tex
@@
 
-We implement this fragmentation by manipulating the iterators used to traverse
-the guest VM's stage-2 page table. In particular, the VM's stage-2 page table
-is a four-level page table, comprising a page global directory (PGD), a
-page upper directory (PUD), a page middle directory (PMD), and multiple page table
-entries (PTEs). TODO (Matthew - Page table fragmentation)
+We implement this fragmentation by manipulating the iterators that provide the next useable
+page at a given level. These iterators are used when traversing the stage-2 page table in order to record
+an additional entry. In the original SeKVM implementation, the pages in each level of the stage-2 page table are
+kept contiguous to one another, and never intermixed with pages from other levels.
+Each iterator starts at a particular offset into the VM's statically allocated stage-2
+page table region, and increments each time a new entry is consumed at that level. The design of these iterators means that they are limited to
+only operating within a large, contiguous region. In particular, the PGD iterator operates over a \textit{16 * PAGE\_SIZE} region, the PUD
+iterator operates over a \textit{SZ\_2M - (16 * PAGE\_SIZE)} region, and the PMD iterator operates over a \textit{3 * SZ\_2M} region.
+
+In contrast, our design eliminates the need for all of these subregions to be contiguous to
+one another and enables fragmentation within each subregion itself. Each iterator can be thought of as a page allocator specific to VM stage-2
+page tables, and operates over a pool of eight, non-contiguous 1MB regions. The FlexAlloc iterators
+first exhaust all pages from a given 1MB region. Once the current 1MB region is exhausted, FlexAlloc
+will jump the iterators to the next 1MB region in the pool. Subsequent traversals of the
+stage-2 page tables with the alloc flag enabled will pull pages from the seperate, second
+non-contiguous 1M region. By removing the requirement for contiguous 8MB regions, FlexAlloc
+enables reliable dynamic allocation of stage-2 page table regions.
+
+We implement two versions of the FlexAlloc iterators. In the first version, we use a single,
+generalized iterator shared across all levels of the stage-2 page table. In the SeKVM implementation,
+the iterators are structured such that a PUD page will never be sandwiches between two PMD pages.
+The first version of the FlexAlloc iterators ignores this semantic and instead is non-discriminatory
+in regards to the particular level a page will be used in. However, we eventually decided to stop development of this
+version before it was fully functional, in favor of version two which more closely resembles recent changes made to SeKVM's
+stage-2 page table allocation scheme.
+
+In the second implementation, we opt to enforce the semantic that pages at each specific level are
+allocated from seperate pools, and thus never intermixed with one another. To accomplish this while still maintaining the same pool of non-contiguous 1MB regions,
+we split the single, general iterator into nine distinct iterators. One iterator is used
+for PUD pages, two iterators are used for PMD pages, and six iterators are used for PTE pages.
+The number of iterators assigned to each level is based on the potential maximum number of
+pages that could be consumed at that level, along with how many distinct 1MB regions each level
+can be allocated from. When a page is allocated at a particular level, the correct iterator is chosen based on
+its assigned level and whether or not its preceding iterators have been exhausted.
+
+The naming semantics of the iterators at each level can be confusing, so we feel it is worth
+clarifying here. The PGD iterator allocates pages to use as entries inside of a PGD page. This
+implies that the PGD iterator, as an example, actually allocates pages that are used in the
+subsequent level, which in this case is the PUD. All iterators follow the pattern of allocating
+pages for a given level to point to. These are actually pages in the next level
+of the stage-2 page tables.
 
@@
 
-TODO (Matthew - DPM)
+To enable dynamic allocation of EL2 memory in general beyond just the scope of
+a VM's stage-2 page tables, we also designed but did not implement
+a dynamic page manager (DPM). DPM is a system that exposes an interface for other ``services'' within EL2 to
+utilize dynamic allocation for their memory regions, while hiding the implementation of actually receiving memory from
+and returning memory to the host. This would be another interesting area of future work.
 

### 2023-12-24 paper.tex
Added iterator diagrams and elaborations in section 2

@@
 \usepackage{filecontents}
+
+\usepackage{graphicx}
+\graphicspath{ {./imgs/} }
 
@@
 \author{
-{\rm Matthew Roger Nelson}\\
+{\rm Matthew Nelson}\\
 Columbia University
@@
 
+Although the metadata structs have a much smaller memory footprint than the stage-2 page tables,
+we designed and partially implemented a fully generalized dynamic allocation interface that
+could be used to support dynamic allocation for all metadata structs which we call the
+Dynamic Page Manager (DPM). DPM handles all of the work required to receive and return
+memory from the hostvisor allowing other "services" in EL2 to exchange memory with DPM using
+a very similiar interface to the alloc_pages and free_pages interface available in EL1. We describe
+this interface in detail in Section 4.
+
 %-------------------------------------------------------------------------------
@@
 As described above, the original implementation of SeKVM statically
-allocates sixteen contiguous 8M regions to the corevisor. When a VM is created,
+allocates sixteen contiguous 8M regions to the corevisor (Figure \ref{fig:sekvmvmregions}). When a VM is created,
 the corevisor uses one of these contiguous 8M regions for the VM's stage-2
@@
 undesirable to continuously allocate huge contiguous regions of memory.
+
+\begin{figure}[h!tbp]
+\caption{SeKVM VM Stage-2 Page Table Region Layout, Statically Allocated}
+\label{fig:sekvmvmregions}
+\centering
+    \includegraphics{sekvm_vm_region}
+\end{figure}
 
@@
 its assigned level and whether or not its preceding iterators have been exhausted.
+
+\begin{figure}[h!tbp]
+\caption{FlexAlloc: First 1MB Region Metadata Setup}
+\label{fig:firstmbregion}
+\centering
+    \includegraphics{firstmbregion}
+\end{figure}
+
+\begin{figure}[h!tbp]
+\caption{FlexAlloc: Second to Eigth 1MB Region Metadata Setup}
+\label{fig:otherregions}
+\centering
+    \includegraphics{otherregions}
+\end{figure}
+
+Each level's iterators have a set number of regions from which to pull additional pages from. Lower
+levels in the table have larger regions than higher levels in the table. When the current iterator
+over a given region is exhausted, the iterator will jump to the next available region and pick up
+where it left off.
+
+\begin{verbatim}
+static u64 inline get_pmd_next(u32 vmid) {
+    ...
+    // pool_index: available iterator
+    // we are using
+    pool_index =
+      el2_data->vm_info[vmid].pte_pool_index;
+    // pool_start: beginning of the region
+    // the pool_index corresponds to
+    pool_start =
+      el2_data->vm_info[vmid].
+      pte_pool_starts[pool_index];
+    // used_pages: number of pages in this
+    // region we have exhausted
+    used_pages = el2_data->vm_info[vmid]
+      .pte_used_pages_vm[pool_index];
+
+    // If we've used up all the pages and we
+    // have another region, then use the new region
+    if (used_pages == PTE_ITER_MAX_PAGES &&
+        pool_index + 1 < PTE_USED_ITER_COUNT) {
+        pool_index += 1;
+        el2_data->vm_info[vmid].
+          pte_pool_index = pool_index;
+
+        // Set the pool_start to the next available
+        // PTE iterator
+        pool_start = el2_data->vm_info[vmid].
+          pte_pool_starts[pool_index];
+        used_pages = el2_data->vm_info[vmid].
+          pte_used_pages_vm[pool_index];
+    }
+
+    // Make sure we have not exhausted all pages
+    // across all regions
+    if (used_pages == PTE_ITER_MAX_PAGES &&
+        pool_index == PTE_USED_ITER_COUNT - 1) {
+        __hyp_panic();
+    }
+
+    // Return the physical address of the next
+    // page that this iterator can allocate
+    return pool_start +
+      (used_pages * PAGE_SIZE);
+}
+\end{verbatim}
 

### 2023-12-24 paper.tex
formatting issues

@@
     if (used_pages == PTE_ITER_MAX_PAGES &&
-        pool_index == PTE_USED_ITER_COUNT - 1) {
+        pool_index == PTE_USED_ITER_COUNT - 1)
+        {
         __hyp_panic();
@@
 We evaluate FlexAlloc by running it on a QEMU-emulated Raspberry Pi 4. In
-particular, we run VMWare Fusion Professional Version 13.0.2 on an M1 Macbook
-Air running MacOS Ventura 13.6. Using VMWare, we create a VM
-with 2 processor cores and 8704 MB memory running Ubuntu 22.04.3 LTS. On this
+particular, we run VMWare Fusion Professional Version 13.0.0 on a 2018 x86-64 Macbook
+Pro running MacOS Ventura 13.6.1. Using VMWare, we create a VM
+with 4 processor cores and 12 GB memory running Ubuntu 22.04.3 LTS. On this
 VM, we run QEMU 6.2.0 to emulate a Raspberry Pi 4. This emulated Raspberry
-Pi 4 serves as our host machine, on which we run FlexAlloc. Our guest VMs
-run vanilla Linux v5.4.
+Pi 4 serves as our host machine, on which we run an SeKVM with the FlexAlloc
+extension. Our guest VMs run vanilla Linux v5.4.
 
@@
 region reclaimed during VM shutdown, and verify that they match each other.
+
+In order to make sure that the host can reasonably be expected to find 8x 1MB
+contiguous regions when booting a VM, we evaluated how many successful
+alloc_pages(1MB) calls could be made before failure. To reproduce our tests,
+simply compile host-sekvm-project with the \textit{ALLOC\_STRESS\_TEST} macro defined.
+Across our three trails, after booting host SeKVM with the FlexAlloc extension,
+we were able to achieve between 7049 and 7054 successful 1MB allocations
+from alloc_pages(1MB) calls. Since each VM only needs 8 successful allocations,
+we believe that the upperbound on the number of VM's due to potential alloc_pages(1MB)
+failures is sufficiently high.
+
+We verified the changes in corevisor's memory footprint by checking the number of
+pages owned by the corevisor according to the \textit{el2\_data\-\>s2\_pages} array at key points
+throughout the VM boot and destruction processes. On a host running SeKVM without the
+FlexAlloc extension, we observed that the Corevisor owned 65549 pages before booting a VM,
+while running a VM, and after destroying a VM. On a host running SeKVM with the FlexAlloc
+extension, we observed that the Corevisor owned 32768 pages before VM boot, 34817 pages
+during the life of the VM, and 32768 pages after destroying the VM. Note that FlexAlloc also
+reduces the static stage2_region size in the linker script to zero as it is no longer needed.
+These findings show that FlexAlloc is able to successfully receive, use, and handback pages
+given by the host for usage exclusively as VM stage-2 page tables.
 
@@
 read from a page table entry that should not actually exist, but that has
-previously been written into the memory region by the hostvisor. Another flaw is that
+previously been written into the memory region by the hostvisor. We also occasionally
+observed stalls when booting subsequent VMs after shutting down other VMs that we
+believe were caused by failing to \textit{memset} the pages to zero. For example, the second
+VM to boot after booting and shutting down the first VM can receive pages with non-zero data
+that, when interpretted as page table entries, lead to undefined behavior.
+
+Another flaw is that
 FlexAlloc does not verify that the eight 1M regions are actually distinct and
@@
 a non-malicious hostvisor, and would require further patches to be fully secure.
+To fix this, FlexAlloc can simply have an added check that verifies all of the
+physical addresses received from the host are unique, 1MB aligned, and non-overlapping.
 

### 2023-12-24 paper.tex
Get paper to compile + Minor figure changes

@@
 \usepackage{amsmath}
+\usepackage{caption}
 
@@
 memory from the hostvisor allowing other "services" in EL2 to exchange memory with DPM using
-a very similiar interface to the alloc_pages and free_pages interface available in EL1. We describe
+a very similiar interface to the alloc\_pages and free\_pages interface available in EL1. We describe
 this interface in detail in Section 4.
@@
 \begin{figure}[h!tbp]
-\caption{SeKVM VM Stage-2 Page Table Region Layout, Statically Allocated}
+\centering
+\captionsetup{justification=centering}
+\includegraphics{sekvm_vm_region}
+\caption{SeKVM: Layout of Statically-Allocated\\Region for VM Stage-2 Page Tables}
 \label{fig:sekvmvmregions}
-\centering
-    \includegraphics{sekvm_vm_region}
 \end{figure}
 
-To get around this limitation, there are two possible approaches. The first approach
+\bigskip To get around this limitation, there are two possible approaches. The first approach
 is to halve the size of the VM's stage-2 page table, so that it can fit within
@@
 \begin{figure}[h!tbp]
-\caption{FlexAlloc: First 1MB Region Metadata Setup}
+\centering
+\captionsetup{justification=centering}
+\includegraphics{firstmbregion}
+\caption{FlexAlloc: Metadata Setup\\for First 1MB Region}
 \label{fig:firstmbregion}
-\centering
-    \includegraphics{firstmbregion}
 \end{figure}
@@
 \begin{figure}[h!tbp]
-\caption{FlexAlloc: Second to Eigth 1MB Region Metadata Setup}
+\centering
+\captionsetup{justification=centering}
+\includegraphics{otherregions}
+\caption{FlexAlloc: Metadata Setup\\for Second to Eigth 1MB Regions}
 \label{fig:otherregions}
-\centering
-    \includegraphics{otherregions}
 \end{figure}
 
-Each level's iterators have a set number of regions from which to pull additional pages from. Lower
+\bigskip Each level's iterators have a set number of regions from which to pull additional pages from. Lower
 levels in the table have larger regions than higher levels in the table. When the current iterator
@@
     // If we've used up all the pages and we
-    // have another region, then use the new region
+    // have another region, then use the new
+    // region.
     if (used_pages == PTE_ITER_MAX_PAGES &&
-        pool_index + 1 < PTE_USED_ITER_COUNT) {
+        pool_index + 1 <
+        PTE_USED_ITER_COUNT) {
         pool_index += 1;
@@
 
-        // Set the pool_start to the next available
-        // PTE iterator
+        // Set the pool_start to the next
+        // available PTE iterator.
         pool_start = el2_data->vm_info[vmid].
@@
 
-    // Make sure we have not exhausted all pages
-    // across all regions
+    // Make sure we have not exhausted all
+    // pages across all regions.
     if (used_pages == PTE_ITER_MAX_PAGES &&
-        pool_index == PTE_USED_ITER_COUNT - 1)
-        {
+            pool_index ==
+            PTE_USED_ITER_COUNT - 1) {
         __hyp_panic();
@@
 
-    // Return the physical address of the next
-    // page that this iterator can allocate
+    // Return the physical address of the
+    // next page that this iterator can
+    // allocate.
     return pool_start +

### 2023-12-24 paper.tex
More formatting changes to compile

@@
 contiguous regions when booting a VM, we evaluated how many successful
-alloc_pages(1MB) calls could be made before failure. To reproduce our tests,
+alloc\_pages(1MB) calls could be made before failure. To reproduce our tests,
 simply compile host-sekvm-project with the \textit{ALLOC\_STRESS\_TEST} macro defined.
@@
 we were able to achieve between 7049 and 7054 successful 1MB allocations
-from alloc_pages(1MB) calls. Since each VM only needs 8 successful allocations,
-we believe that the upperbound on the number of VM's due to potential alloc_pages(1MB)
+from alloc\_pages(1MB) calls. Since each VM only needs 8 successful allocations,
+we believe that the upperbound on the number of VM's due to potential alloc\_pages(1MB)
 failures is sufficiently high.
@@
 during the life of the VM, and 32768 pages after destroying the VM. Note that FlexAlloc also
-reduces the static stage2_region size in the linker script to zero as it is no longer needed.
+reduces the static stage2\_region size in the linker script to zero as it is no longer needed.
 These findings show that FlexAlloc is able to successfully receive, use, and handback pages

### 2023-12-24 paper.tex
Somewhat resolved formatting issues, still broken

--- 2023-12-24 paper.tex
@@
 memory from the hostvisor allowing other "services" in EL2 to exchange memory with DPM using
-a very similiar interface to the alloc_pages and free_pages interface available in EL1. We describe
+a very similiar interface to the \textit{alloc\_pages} and \textit{free\_pages} interface available in EL1. We describe
 this interface in detail in Section 4.
@@
 contiguous regions when booting a VM, we evaluated how many successful
-alloc_pages(1MB) calls could be made before failure. To reproduce our tests,
+\textit{alloc\_pages(1MB)} calls could be made before failure. To reproduce our tests,
 simply compile host-sekvm-project with the \textit{ALLOC\_STRESS\_TEST} macro defined.
@@
 we were able to achieve between 7049 and 7054 successful 1MB allocations
-from alloc_pages(1MB) calls. Since each VM only needs 8 successful allocations,
-we believe that the upperbound on the number of VM's due to potential alloc_pages(1MB)
+from \textit{alloc\_pages(1MB)} calls. Since each VM only needs 8 successful allocations,
+we believe that the upperbound on the number of VM's due to potential \textit{alloc\_pages(1MB)}
 failures is sufficiently high.
@@
 We verified the changes in corevisor's memory footprint by checking the number of
-pages owned by the corevisor according to the \textit{el2\_data\-\>s2\_pages} array at key points
+pages owned by the corevisor according to the \textit{s2\_pages} array in \textit{el2\_data} at key points
 throughout the VM boot and destruction processes. On a host running SeKVM without the

### 2023-12-24 paper.tex

--- 2023-12-24 paper.tex
@@
 memory from the hostvisor allowing other "services" in EL2 to exchange memory with DPM using
-a very similiar interface to the alloc\_pages and free\_pages interface available in EL1. We describe
+a very similiar interface to the \textit{alloc\\_pages} and \textit{free\\_pages} interface available in EL1. We describe
 this interface in detail in Section 4.
@@
 contiguous regions when booting a VM, we evaluated how many successful
-alloc\_pages(1MB) calls could be made before failure. To reproduce our tests,
+\textit{alloc\_pages(1MB)} calls could be made before failure. To reproduce our tests,
 simply compile host-sekvm-project with the \textit{ALLOC\_STRESS\_TEST} macro defined.
@@
 we were able to achieve between 7049 and 7054 successful 1MB allocations
-from alloc\_pages(1MB) calls. Since each VM only needs 8 successful allocations,
-we believe that the upperbound on the number of VM's due to potential alloc\_pages(1MB)
+from \textit{alloc\_pages(1MB)} calls. Since each VM only needs 8 successful allocations,
+we believe that the upperbound on the number of VM's due to potential \textit{alloc\_pages(1MB)}
 failures is sufficiently high.

### 2023-12-24 paper.tex
fixed format issues

@@
 memory from the hostvisor allowing other "services" in EL2 to exchange memory with DPM using
-a very similiar interface to the \textit{alloc\\_pages} and \textit{free\\_pages} interface available in EL1. We describe
+a very similiar interface to the \textit{alloc\_pages} and \textit{free\_pages} interface available in EL1. We describe
 this interface in detail in Section 4.

### 2023-12-24 paper.tex
added evaluation edits

@@
 Another key shortcoming of our design is that it does not fully implement
-dynamic memory partitioning between the corevisor and hostvisor. As mentioned in
+dynamic memory partitioning between the corevisor and hostvisor outside of stage-2 page tables for VMs. As mentioned in
 the introduction, another area in which corevisor memory scales with the number
@@
 \textit{struct el2\_vm\_info} object is only 768 bytes, this would not waste a
-significant amount of memory even if fewer than sixteen VMs are created.
+significant amount of memory even if fewer than sixteen VMs are created. Regarding the
+\textit{shared\_data\_region}, since each VM requires a \textit{struct kvm}, the size of
+this region would either need to be dynamically allocated or the static size would
+need to be increases in order to lift the 16 VM cap present in SeKVM. However, since the static region
+size has been decreased by 128 MB within FlexAlloc and the size of \textit{struct kvm}'s is relatively
+small, inflating this region to accomodate 32+ VMs would still result in substantial memory savings
+thanks to FlexAlloc's changes. In the Future Work section, we propose a potential solution to
+the issue of bringing dynamic allocation to EL2 metadata structures that are not page aligned
+which we call DPM.
 
@@
 the stage-2 page table for a new VM. We could perhaps delay this limit by breaking
-the VM stage-2 page tables down into even smaller regions of memory, but this would
-likely lead to more complicated logic when traversing the page table. In addition, we do not verify whether fragmenting the stage-2 page table results
-in any degradation in VM performance. Ideally, we would run a memory-intensive
+the VM stage-2 page tables down into even smaller regions of memory. In addition, we do not verify whether fragmenting the stage-2 page table results
+in any degradation in VM performance. Given the less-than-ideal test environment
+we have at our disposal, we did not investigate performance metrics related to execution time
+for our implementation and only considered memory footprint metrics. Ideally, we would run a memory-intensive
 benchmark on our guest VMs to verify that our iterators work properly under a
-high memory load. Finally, we do not quantify the change in corevisor memory usage.
-We assume that by not allocating the memory for VM stage-2 page tables at bootup,
-corevisor memory will automatically be reduced by an amount matching the
-size of this static allocation. This should be verified using actual data. For
-instance, we could periodically iterate over every page and see how many pages
-are owned by the corevisor before and after VM creation.
+high memory load across multiple VMs.
 
@@
 into eight distinct 1M pages, because \textit{alloc\_pages} does not allow us
-to allocate contiguous regions larger than 4M. It might be possile to avoid this
+to allocate contiguous regions larger than 4M. We also wanted to be reasonably certain
+that the size of our regions would not impeed the host's ability to launch VMs. It might be possile to avoid this
 by looking at other Linux memory allocators explicitly designed for the purpose
-of large contiguous allocations.
+of large contiguous allocations and adding such an allocator to the host.
 
+\subsection{Dynamic Page Manager}
 To enable dynamic allocation of EL2 memory in general beyond just the scope of
-a VM's stage-2 page tables, we also designed but did not implement
+a VM's stage-2 page tables, we also designed but did not fully implement
 a dynamic page manager (DPM). DPM is a system that exposes an interface for other ``services'' within EL2 to

### 2023-12-24 paper.tex
most of dpm done

@@
 
-\subsection{Dynamic Page Manager}
+\subsection{Dynamic Page Manager - Overview}
 To enable dynamic allocation of EL2 memory in general beyond just the scope of
@@
 utilize dynamic allocation for their memory regions, while hiding the implementation of actually receiving memory from
-and returning memory to the host. This would be another interesting area of future work.
+and returning memory to the host. This would be another interesting area of future work. Please
+see the \textit{dpm} branch of the linked repository for a partial DPM implementation.
+
+By implementing a means by which EL2 can dynamically “request” and “hand back” contiguous
+regions of pages from EL1 arbitrarily, EL2 would no longer be limited to using statically allocated, fixed size regions.
+We propose the following Dynamic Page Manager (DPM) interface for EL1 and EL2 to interact over
+when dynamically allocating memory for EL2 to use arbitrarily.
+
+Other services within EL2 can use DPM to dynamically request contiguous regions of pages.
+These services will be responsible for managing the memory contents of these pages once DPM
+allocates the region to that specific service. The service will be required to signal DPM
+when that contiguous region is no longer in use which is similar to the kmalloc and kfree model.
+
+DPM operates at the contiguous region granularity rather than the page granularity because
+a lot of the existing SeKVM code assumes contiguous regions. If DPM were to instead piece
+together pages to form fragmented regions, all of this other code would break. Thus DPM
+can actually be thought of as a region manager rather than a page manager. DPM is structured
+in such a way that a second page-granularity version of DPM could operate alongside the
+region-granularity version of DPM thus allowing some services to remaining dependent on
+contiguous regions while others adapt to use separated regions in support of better memory
+utilization by EL2.
+
+\subsection{Dynamic Page Manager - Interface}
+Additional hypercalls accessible by EL1 host to DPM:
+
+\begin{itemize}
+\item Giving allocated pages from EL1 to EL2:
+\textit{hvc\_give\_pages(starting\_phys\_addr, page\_aligned region)}
+\item Reclaim unused pages from EL2 back to EL1:
+\textit{hvc\_reclaim\_pages(output\_struct\_array\_page\_starting\_phys\_addr)}
+All regions being handed back to the host are recorded in the return page which is
+passed by the host. Every physical address that EL2 records in this page is considered
+freed and no longer managed by DPM.
+\end{itemize}
+
+DPM Interface for usage internally by EL2 services:
+\begin{itemize}
+\item Requesting a contiguous region of a given size from the pool:
+\textit{dpm\_alloc\_region(region size)}:
+	Returns an address to the region if one was available in the pool,
+    and 0 to indicate error which the requesting service must propagate back to EL1
+\item
+\textit{dpm\_free\_region(starting\_region\_addr, optional: size)}:
+Returns a previously allocated region back to the pool from the EL2 service that used it
+\end{itemize}
+
+\subsection{Dynamic Page Manager - Metadata}
+\begin{verbatim}
+// DPM Overall Metadata
+struct dpm { // contained in struct el2_data
+    u64 base_page_addr; // Address to the first
+                        // dpm_region page
+    u64 region_count;   // number of regions currently
+                        // in the pool
+    u64 page_count;     // number of pages currently
+                        // in the pool
+};
+
+// DPM Region-specific Metadata
+struct dpm_region {
+    u64 start_addr; // addr to beginning of first
+                    // page in the region
+    u64 page_count; // number of pages in the region
+
+    u8 flags; // DXXX XXXA, where A is allocated flag,
+              // D is DPM flag meaning that region
+              // holds DPM metadata, X is available bit
+
+    struct dpm_region* prev; // consider changing
+                             // these to u64 physical addresses
+    struct dpm_region* next;
+};
+\end{verbatim}
+
+DPM manages regions of pages in form of a circular linked list. When an EL2 service requests
+a region of a specific size from DPM, DPM will search this list for an available match.
+If no matches are found, DPM returns an error indicating to the EL2 service that it must
+propogate the error back to the host. Upon receiving the error, the host will make an
+additional hypercall that passes a new region of some size to DPM. After restarting the original
+hypercall, DPM will successfully find a region to hand to the EL2 service.
+
+If the host is under a lot of memory pressure, the host can request that DPM give control
+of available regions back to the host. The host passes a page that DPM will populate with
+an array of metadata pertaining to regions that it is handing back. DPM will first unmap this
+page from the host. DPM will then search through the list for all available regions and add them
+to the page. DPM will finally remap the return page to the host and return. The host can then
+call \textit{free_pages} on all of the return regions.
+
+In order for DPM to receive pages that it needs for it's own metadata, if the host receives an
+error when trying to hand a region of some size to DPM, then the host will know that it needs
+to pass a region of size PAGE_SIZE first to DPM in a seperate hypercall and then retry the
+original region pass hypercall before finally retrying the original hypercall pertaining
+to some other EL2 service. When DPM receives a region and notices that it is out of metadata
+space, DPM will search for a region of PAGE_SIZE to claim for itself before processing
+the larger region. If one is found, then DPM will claim it and process the region hand off
+as normal.
+
+\subsection{Dynamic Page Manager - Considerations and Drawbacks}
 

### 2023-12-24 paper.tex
DPM content done

@@
 
-    struct dpm_region* prev; // consider changing
-                             // these to u64 physical addresses
+    struct dpm_region* prev;
     struct dpm_region* next;
@@
 to the page. DPM will finally remap the return page to the host and return. The host can then
-call \textit{free_pages} on all of the return regions.
+call \textit{free\_pages} on all of the return regions.
 
@@
 error when trying to hand a region of some size to DPM, then the host will know that it needs
-to pass a region of size PAGE_SIZE first to DPM in a seperate hypercall and then retry the
+to pass a region of size \textit{PAGE\_SIZE} first to DPM in a seperate hypercall and then retry the
 original region pass hypercall before finally retrying the original hypercall pertaining
 to some other EL2 service. When DPM receives a region and notices that it is out of metadata
-space, DPM will search for a region of PAGE_SIZE to claim for itself before processing
+space, DPM will search for a region of \textit{PAGE\_SIZE} to claim for itself before processing
 the larger region. If one is found, then DPM will claim it and process the region hand off
@@
 \subsection{Dynamic Page Manager - Considerations and Drawbacks}
+DPM allows for arbitrary request of regions from EL2 which is valuable when we may or may not
+need additional pages. If we know some hypercall will require EL2 to receive more pages,
+the error-check-retry pathway may be an unnecessary cost. For example, when we want to
+boot a new VM, we know EL2 will need a stage 2 region for the VM. However, since DPM allows
+regions to be reused after they are made available again, a new VM launch can reuse the
+region from a previously shutdown VM. DPM is designed make it convenient to program against,
+but it’s generalized strategy can lead to extra, avoidable work. We designed DPM with
+programmability in mind over efficiency in hopes that it would make full dynamic allocation
+attainable for SeKVM services by avoiding excessively customized hand offs between the host
+and corevisor. Although we only partially implemented DPM, it would be interesting to
+compare the real-world performance metrics between our current stage-2 dynamic allocation
+strategy and a stage-2 dynamic allocation strategy built on top of DPM and a
+check-error-retry methodology.
 

### 2023-12-24 paper.tex
Standardize terminology: Pools, Entries, Iterators, Allocations, Pages

@@
 
-Although the metadata structs have a much smaller memory footprint than the stage-2 page tables,
-we designed and partially implemented a fully generalized dynamic allocation interface that
-could be used to support dynamic allocation for all metadata structs which we call the
-Dynamic Page Manager (DPM). DPM handles all of the work required to receive and return
-memory from the hostvisor allowing other "services" in EL2 to exchange memory with DPM using
+Although the \textit{struct el2\_vm\_info} objects have a much smaller memory footprint than the stage-2 page tables,
+we designed and partially implemented a fully-generalized dynamic allocation interface that
+could be used to support dynamic allocation for all metadata structs. We call this the
+Dynamic Page Manager (DPM). DPM handles all of the work required to receive memory from and return
+memory to the hostvisor, allowing other ``services'' in EL2 to exchange memory with the DPM using
 a very similiar interface to the \textit{alloc\_pages} and \textit{free\_pages} interface available in EL1. We describe
-this interface in detail in Section 4.
+this interface in detail in Section 4. However, the next section focuses
+exclusively on the changes made to support dynamic allocation of the memory
+needed for VM stage-2 page tables.
 
@@
 page table. To implement a system of dynamic allocation, we want the hostvisor
-to dynamically reserve a contiguous 8M region at runtime when a VM is created, and then
+to dynamically reserve an 8M region at runtime when a VM is created, and then
 pass this memory region to the corevisor. The kernel macro \textit{alloc\_pages}
@@
 We implement this fragmentation by manipulating the iterators that provide the next useable
-page at a given level. These iterators are used when traversing the stage-2 page table in order to record
-an additional entry. In the original SeKVM implementation, the pages in each level of the stage-2 page table are
-kept contiguous to one another, and never intermixed with pages from other levels.
-Each iterator starts at a particular offset into the VM's statically allocated stage-2
-page table region, and increments each time a new entry is consumed at that level. The design of these iterators means that they are limited to
-only operating within a large, contiguous region. In particular, the PGD iterator operates over a \textit{16 * PAGE\_SIZE} region, the PUD
-iterator operates over a \textit{SZ\_2M - (16 * PAGE\_SIZE)} region, and the PMD iterator operates over a \textit{3 * SZ\_2M} region.
+page for entries at a given level of the page table. In the original SeKVM implementation, the pages containing entries from some level of the stage-2 page table are
+kept contiguous to one another, and are never intermixed with pages containing entries from other levels. Hence, there are distinct iterators for each level.
+Each iterator starts at a particular offset into the VM's statically-allocated 8M stage-2
+page table region, and increments each time we need a new page for a new entry at that level. In particular, the PGD iterator operates over a \textit{16 * PAGE\_SIZE} pool, the PUD
+iterator operates over a \textit{SZ\_2M - (16 * PAGE\_SIZE)} pool, and the PMD iterator operates over a \textit{3 * SZ\_2M} pool.
+Each of these pools must be contiguous, and the three pools must be contiguous relative to one another.
 
-In contrast, our design eliminates the need for all of these subregions to be contiguous to
-one another and enables fragmentation within each subregion itself. Each iterator can be thought of as a page allocator specific to VM stage-2
-page tables, and operates over a pool of eight, non-contiguous 1MB regions. The FlexAlloc iterators
-first exhaust all pages from a given 1MB region. Once the current 1MB region is exhausted, FlexAlloc
-will jump the iterators to the next 1MB region in the pool. Subsequent traversals of the
-stage-2 page tables with the alloc flag enabled will pull pages from the seperate, second
-non-contiguous 1M region. By removing the requirement for contiguous 8MB regions, FlexAlloc
-enables reliable dynamic allocation of stage-2 page table regions.
+The naming semantics of the iterators at each level can be confusing, so we feel they are worth
+clarifying here. The iterator for level $n$ allocates pages for entries in level $n$ to point to.
+Hence, the iterator for level $n$ points to the pool used to allocate
+entries belonging to level $n-1$. For instance, the PGD iterator allocates pages
+for PGD entries to point to. In other words, the PGD iterator actually allocates
+pages used to hold PUD entries. Hence, the PGD iterator is initialized to point to the start of the PUD pool.
 
-We implement two versions of the FlexAlloc iterators. In the first version, we use a single,
-generalized iterator shared across all levels of the stage-2 page table. In the SeKVM implementation,
-the iterators are structured such that a PUD page will never be sandwiches between two PMD pages.
-The first version of the FlexAlloc iterators ignores this semantic and instead is non-discriminatory
-in regards to the particular level a page will be used in. However, we eventually decided to stop development of this
+Our design eliminates the need for all of these pools to be contiguous to
+one another, and enables fragmentation within each pool itself. To achieve this goal,
+we experiment with two distinct implementations. As mentioned above, the SeKVM implementation
+structures iterators such that a page with PUD entries will never be sandwiched between a page with PMD entries.
+In our first implementation, we ignore this semantic. In particular, we create a single,
+generalized iterator shared across all levels of the stage-2 page table. This
+iterator traverses all eight 1M regions. When it exhausts the space in a 1M region, it moves
+to the start of the next 1M region. However, we eventually decided to stop development of this
 version before it was fully functional, in favor of version two which more closely resembles recent changes made to SeKVM's
@@
 
-In the second implementation, we opt to enforce the semantic that pages at each specific level are
-allocated from seperate pools, and thus never intermixed with one another. To accomplish this while still maintaining the same pool of non-contiguous 1MB regions,
-we split the single, general iterator into nine distinct iterators. One iterator is used
-for PUD pages, two iterators are used for PMD pages, and six iterators are used for PTE pages.
-The number of iterators assigned to each level is based on the potential maximum number of
-pages that could be consumed at that level, along with how many distinct 1MB regions each level
-can be allocated from. When a page is allocated at a particular level, the correct iterator is chosen based on
-its assigned level and whether or not its preceding iterators have been exhausted.
+In our second implementation, we enforce the semantic that pages at each specific level are
+allocated from seperate pools, and thus are never intermixed with one another. To accomplish
+this while still maintaining the same pool of non-contiguous 1MB regions,
+we split the single, generalized iterator into nine distinct iterators. 
+One PGD iterator is used to allocate pages from the PUD pool, two PUD iterators are used to allocate pages from the PMD pool,
+and six PMD iterators are used to allocate pages from the PTE pool. There are two key things to
+note here. First, we retain the existing scheme partitioning the stage-2 page table into
+three distinct pools: a PUD pool, a PMD pool, and a PTE pool.
+For instance, the PUD pool is still sized at \textit{16 * PAGE\_SIZE}. The
+second key thing to note is that our iterators never cross the boundary
+between two distinct 1M regions. Instead, if a pool needs to be split across
+two or more 1M regions, there is one distinct iterator for each 1M region
+covered by the pool. As a result, each of these iterators covers a contiguous subregion of the pool.
+Allocations from that pool begin by using the first iterator. When the first
+subregion of that pool is exhausted, allocations move on to use the next iterator.
+Essentially, when a page is allocated from a given pool, the correct iterator is chosen based on
+the pool in question and whether or not the preceding iterators for that pool have been exhausted.
 
+% We do that by modfifying , the FlexAlloc
+% iterators operate over a pool of eight non-contiguous 1MB regions. The FlexAlloc iterators
+% first exhaust all entries from a given 1MB region. Once the current 1MB region is exhausted, FlexAlloc
+% will jump the iterators to the next 1MB region in the pool. Subsequent traversals of the
+% stage-2 page tables with the alloc flag enabled will pull pages from the seperate, second
+% non-contiguous 1M region. By removing the requirement for contiguous 8MB regions, FlexAlloc
+% enables reliable dynamic allocation of stage-2 page table regions.
+
+\bigskip
 \begin{figure}[h!tbp]
@@
 \includegraphics{otherregions}
-\caption{FlexAlloc: Metadata Setup\\for Second to Eigth 1MB Regions}
+\caption{FlexAlloc: Metadata Setup\\for Second to Eighth 1MB Regions}
 \label{fig:otherregions}
@@
 
-\bigskip Each level's iterators have a set number of regions from which to pull additional pages from. Lower
-levels in the table have larger regions than higher levels in the table. When the current iterator
-over a given region is exhausted, the iterator will jump to the next available region and pick up
-where it left off.
+In particular, the PUD pool is in the middle of the first 1M region, and there is a single PGD iterator
+pointing to the start of that pool (Figure \ref{fig:firstmbregion}). The PMD pool consumes the end of the first 1M region
+and the entire second 1M region. Hence, the first PUD iterator points to an address in
+the first 1M region, and the second PUD Iterator points to the start of the
+second 1M region (Figures \ref{fig:firstmbregion}, \ref{fig:otherregions}). Finally, the PTE pool consumes the third to eighth 1M regions.
+As a result, there is a distinct PMD iterator for each of these six regions
+(Figure \ref{fig:otherregions}).
+
+The code below illustrates how allocations from a given pool are made using
+the first possible iterator whose pool subregion has not
+been exhausted:
 
@@
 \end{verbatim}
-
-The naming semantics of the iterators at each level can be confusing, so we feel it is worth
-clarifying here. The PGD iterator allocates pages to use as entries inside of a PGD page. This
-implies that the PGD iterator, as an example, actually allocates pages that are used in the
-subsequent level, which in this case is the PUD. All iterators follow the pattern of allocating
-pages for a given level to point to. These are actually pages in the next level
-of the stage-2 page tables.
 

### 2023-12-24 paper.tex
Some changes to eval, e.g. fix mistake re: STAGE2 region, standardize tenses

@@
 As a result, there is a distinct PMD iterator for each of these six regions
-(Figure \ref{fig:otherregions}).
-
-The code below illustrates how allocations from a given pool are made using
+(Figure \ref{fig:otherregions}). The code below illustrates how allocations from a given pool are made using
 the first possible iterator whose pool subregion has not
@@
 
-In order to make sure that the host can reasonably be expected to find 8x 1MB
-contiguous regions when booting a VM, we evaluated how many successful
-\textit{alloc\_pages(1MB)} calls could be made before failure. To reproduce our tests,
-simply compile host-sekvm-project with the \textit{ALLOC\_STRESS\_TEST} macro defined.
-Across our three trails, after booting host SeKVM with the FlexAlloc extension,
-we were able to achieve between 7049 and 7054 successful 1MB allocations
-from \textit{alloc\_pages(1MB)} calls. Since each VM only needs 8 successful allocations,
-we believe that the upperbound on the number of VM's due to potential \textit{alloc\_pages(1MB)}
+In order to verify that the hostvisor will generally be able to find eight distinct contiguous 1M regions when booting a VM, we evaluate how many times we can successfully call
+\textit{alloc\_pages(1M)} before a null pointer is returned. To reproduce our tests,
+simply compile \textit{host-sekvm-project} with the \textit{ALLOC\_STRESS\_TEST} macro defined.
+Across our three trials, after booting host SeKVM with the FlexAlloc extension,
+we were able to achieve between 7049 and 7054 successful 1M allocations
+from calls to \textit{alloc\_pages(1M)}. Since each VM only needs eight successful allocations for its stage-2 page table,
+we believe that the upper bound on the number of VMs we can create before running into \textit{alloc\_pages(1M)}
 failures is sufficiently high.
 
-We verified the changes in corevisor's memory footprint by checking the number of
-pages owned by the corevisor according to the \textit{s2\_pages} array in \textit{el2\_data} at key points
-throughout the VM boot and destruction processes. On a host running SeKVM without the
-FlexAlloc extension, we observed that the Corevisor owned 65549 pages before booting a VM,
+We also verify the changes in the corevisor's memory footprint by checking the number of
+pages owned by the corevisor, using the \textit{s2\_pages} array in \textit{el2\_data} at key points
+when a VM is created and destroyed. On a host running SeKVM without the
+FlexAlloc extension, we observe that the corevisor owns 65549 pages before booting a VM,
 while running a VM, and after destroying a VM. On a host running SeKVM with the FlexAlloc
-extension, we observed that the Corevisor owned 32768 pages before VM boot, 34817 pages
-during the life of the VM, and 32768 pages after destroying the VM. Note that FlexAlloc also
-reduces the static stage2\_region size in the linker script to zero as it is no longer needed.
-These findings show that FlexAlloc is able to successfully receive, use, and handback pages
-given by the host for usage exclusively as VM stage-2 page tables.
+extension, we observe that the Corevisor owns 32768 pages before VM boot, 34817 pages
+during the life of the VM, and 32768 pages after destroying the VM.
+Note that FlexAlloc also drastically reduces the size of the statically-allocated \textit{STAGE2\_PAGES\_SIZE} region in the linker script,
+since it now only needs to contain the stage-2 page tables of the hostvisor and the corevisor.
+These findings show that the corevisor in FlexAlloc is able to successfully receive, use, and return pages
+given to it by the hostvisor for constructing a VM's stage-2 page table.
+
+%
 

### 2023-12-24 paper.tex
dpm fully complete
Final cleanup

@@
 Dynamic Page Manager (DPM). DPM handles all of the work required to receive memory from and return
-memory to the hostvisor, allowing other ``services'' in EL2 to exchange memory with the DPM using
+memory to the hostvisor, allowing other ``services'' in EL2 to exchange memory with DPM using
 a very similiar interface to the \textit{alloc\_pages} and \textit{free\_pages} interface available in EL1. We describe
@@
 
-%
-
 %-------------------------------------------------------------------------------
@@
 previously been written into the memory region by the hostvisor. We also occasionally
-observed stalls when booting subsequent VMs after shutting down other VMs that we
-believe were caused by failing to \textit{memset} the pages to zero. For example, the second
-VM to boot after booting and shutting down the first VM can receive pages with non-zero data
-that, when interpretted as page table entries, lead to undefined behavior.
+observed stalls when shutting down VMs and booting new ones, which we
+believe is caused by our failure to \textit{memset} the pages to zero. For example,
+if we shut down a VM and boot up a new one, the second VM's stage-2 page table could contain pages previously
+used by the first VM's stage-2 page table. These pages could contain non-zero data,
+leading to undefined behavior when interpreted as page table entries.
 
@@
 a non-malicious hostvisor, and would require further patches to be fully secure.
-To fix this, FlexAlloc can simply have an added check that verifies all of the
-physical addresses received from the host are unique, 1MB aligned, and non-overlapping.
+To fix this, FlexAlloc could simply have an additional check verifying that all of the
+physical addresses received from the host are unique, 1M-aligned, and non-overlapping.
 
-Another key shortcoming of our design is that it does not fully implement
-dynamic memory partitioning between the corevisor and hostvisor outside of stage-2 page tables for VMs. As mentioned in
+Finally, our design does not fully implement
+dynamic memory partitioning between the corevisor and hostvisor, outside of VM stage-2 page tables. As mentioned in
 the introduction, another area in which corevisor memory scales with the number
@@
 \textit{struct el2\_vm\_info} object is only 768 bytes, this would not waste a
-significant amount of memory even if fewer than sixteen VMs are created. Regarding the
-\textit{shared\_data\_region}, since each VM requires a \textit{struct kvm}, the size of
-this region would either need to be dynamically allocated or the static size would
-need to be increases in order to lift the 16 VM cap present in SeKVM. However, since the static region
-size has been decreased by 128 MB within FlexAlloc and the size of \textit{struct kvm}'s is relatively
-small, inflating this region to accomodate 32+ VMs would still result in substantial memory savings
-thanks to FlexAlloc's changes. In the Future Work section, we propose a potential solution to
-the issue of bringing dynamic allocation to EL2 metadata structures that are not page aligned
-which we call DPM.
+significant amount of memory even if fewer than sixteen VMs are created. Moreover,
+each VM also requires a \textit{struct kvm}. Hence, to lift the sixteen-VM cap present
+in SeKVM, we would need to either dynamically allocate the memory for this struct,
+or increase the size of the statically-allocated region for VM metadata.
+The latter approach is probably reasonable.
+Since the changes to the VM stage-2 page tables already decrease the size
+of the statically-allocated region by 128 MB, re-inflating this region to accomodate 32+ VMs would still result in substantial memory savings. 
+The former approach is possible as well. In the subsection on Future Work, we propose DPM as a potential solution to
+the issue of bringing dynamic allocation to EL2 metadata structures that are not page-aligned.
 
@@
 We also do not fully evaluate our implementation. For one, we do not stress-test our implementation by running a large number
-of VMs. To do this, we would have to modify the macros used to define the current
-sixteen-VM limit. It is entirely conceivable that the hostvisor might not be able
-to allocate 1M contiguous regions of memory past a certain point if memory
-pressure and external fragmentation are high. In this case, the hostvisor would
-not be able to allocate the memory regions needed for the corevisor to create
-the stage-2 page table for a new VM. We could perhaps delay this limit by breaking
-the VM stage-2 page tables down into even smaller regions of memory. In addition, we do not verify whether fragmenting the stage-2 page table results
-in any degradation in VM performance. Given the less-than-ideal test environment
-we have at our disposal, we did not investigate performance metrics related to execution time
-for our implementation and only considered memory footprint metrics. Ideally, we would run a memory-intensive
+of VMs. In addition, we do not verify whether fragmenting the stage-2 page table results
+in any degradation in VM performance. Given the limitations of our test environment, we
+only considered metrics related to the corevisor's memory footprint, and did not investigate performance metrics related to execution time. Ideally, we would run a memory-intensive
 benchmark on our guest VMs to verify that our iterators work properly under a
@@
 
-\subsection{Dynamic Page Manager - Overview}
+\subsection{Dynamic Page Manager}
+
+\subsubsection{Overview}
 To enable dynamic allocation of EL2 memory in general beyond just the scope of
@@
 
-By implementing a means by which EL2 can dynamically “request” and “hand back” contiguous
-regions of pages from EL1 arbitrarily, EL2 would no longer be limited to using statically allocated, fixed size regions.
-We propose the following Dynamic Page Manager (DPM) interface for EL1 and EL2 to interact over
-when dynamically allocating memory for EL2 to use arbitrarily.
-
-Other services within EL2 can use DPM to dynamically request contiguous regions of pages.
-These services will be responsible for managing the memory contents of these pages once DPM
-allocates the region to that specific service. The service will be required to signal DPM
-when that contiguous region is no longer in use which is similar to the kmalloc and kfree model.
+By implementing an interface for EL2 to dynamically ``request'' and ``return'' arbitrary
+regions of contiguous memory from EL1, the corevisor would no longer be limited to using statically-allocated, fixed-size regions.
+Instead, other services within EL2 could use DPM to dynamically request regions of contiguous memory.
+These services would be responsible for managing the contents of these pages once DPM
+allocates these pages to that specific service. The service would be required to signal DPM
+when that contiguous region is no longer in use, using an interface that is similar to the kmalloc and kfree model.
 
@@
 together pages to form fragmented regions, all of this other code would break. Thus DPM
-can actually be thought of as a region manager rather than a page manager. DPM is structured
+can actually be thought of as a memory-region manager rather than a page manager. DPM is structured
 in such a way that a second page-granularity version of DPM could operate alongside the
-region-granularity version of DPM thus allowing some services to remaining dependent on
-contiguous regions while others adapt to use separated regions in support of better memory
+region-granularity version of DPM, thus allowing some services to remain dependent on
+contiguous regions while others adapt to use separated non-contiguous regions, in order to achieve better memory
 utilization by EL2.
 
-\subsection{Dynamic Page Manager - Interface}
-Additional hypercalls accessible by EL1 host to DPM:
+\subsubsection{Interface}
+Our design of DPM exposes the following additional hypercalls to the
+hostvisor running at EL1:
 
 \begin{itemize}
-\item Giving allocated pages from EL1 to EL2:
-\textit{hvc\_give\_pages(starting\_phys\_addr, page\_aligned region)}
-\item Reclaim unused pages from EL2 back to EL1:
-\textit{hvc\_reclaim\_pages(output\_struct\_array\_page\_starting\_phys\_addr)}
-All regions being handed back to the host are recorded in the return page which is
-passed by the host. Every physical address that EL2 records in this page is considered
+\item Allocating pages to EL2:\\
+\textit{hvc\_give\_pages(base\_phys\_addr\_of\_region, size\_of\_region)}.
+Note that the size must be page-aligned.
+\item Reclaiming pages from EL2:\\
+\textit{hvc\_reclaim\_pages(base\_phys\_addr\_of\_output\_struct)}.
+The output struct is provided by the hostvisor, and should contain an array of
+physical addresses. Services at EL2 should record the physical addresses
+of  regions being handed back to the hostvisor in the output
+struct. The hostvisor can then read from the output struct and free the relevant pages.
+Every physical address that EL2 records in this output struct is considered
 freed and no longer managed by DPM.
@@
 
-DPM Interface for usage internally by EL2 services:
+\noindent Our design of DPM also exposes the following interface to services running
+at EL2 that want dynamically-allocated memory:
 \begin{itemize}
-\item Requesting a contiguous region of a given size from the pool:
-\textit{dpm\_alloc\_region(region size)}:
+\item Requesting a contiguous region of a given size from the DPM pool:
+\textit{dpm\_alloc\_region(size)}.
 	Returns an address to the region if one was available in the pool,
-    and 0 to indicate error which the requesting service must propagate back to EL1
+    and 0 to indicate an error which the requesting service must propagate back to EL1.
 \item
 \textit{dpm\_free\_region(starting\_region\_addr, optional: size)}:
-Returns a previously allocated region back to the pool from the EL2 service that used it
+Signals that a previously-allocated region can be returned to the DPM pool.
 \end{itemize}
 
-\subsection{Dynamic Page Manager - Metadata}
+\subsubsection{Metadata}
+
+DPM manages regions of pages using a circular linked list (Figure \ref{fig:dpmmetadata}). When an EL2 service requests
+a region of a specific size from DPM, DPM will search this list for an available match.
+If no matches are found, DPM returns an error indicating to the EL2 service that it must
+propogate the error back to the host. Upon receiving the error, the host will make an
+additional hypercall that passes a new region of some size to DPM. After restarting the original
+hypercall, DPM will successfully find a region to hand to the EL2 service. The code
+below illustrates the data structures used to store this metadata.
+
 \begin{verbatim}
 // DPM Overall Metadata
-struct dpm { // contained in struct el2_data
-    u64 base_page_addr; // Address to the first
+// Contained in struct el2_data
+struct dpm {
+    // Address to the first
     // dpm_region page
-    u64 region_count;   // number of regions currently
-                        // in the pool
-    u64 page_count;     // number of pages currently
-                        // in the pool
+    u64 base_page_addr;
+
+    // Number of regions currently in pool
+    u64 region_count;
+
+    // Number of pages currently in pool
+    u64 page_count;     
 };
 
-// DPM Region-specific Metadata
+// DPM Region-Specific Metadata
 struct dpm_region {
-    u64 start_addr; // addr to beginning of first
+    // Addr of the beginning of the first
     // page in the region
-    u64 page_count; // number of pages in the region
+    u64 start_addr;
 
-    u8 flags; // DXXX XXXA, where A is allocated flag,
+    // Number of pages in region
+    u64 page_count;
+
+    // DXXX XXXA, where A is allocated flag,
     // D is DPM flag meaning that region
     // holds DPM metadata, X is available bit
+    u8 flags;
 
@@
 
-DPM manages regions of pages in form of a circular linked list. When an EL2 service requests
-a region of a specific size from DPM, DPM will search this list for an available match.
-If no matches are found, DPM returns an error indicating to the EL2 service that it must
-propogate the error back to the host. Upon receiving the error, the host will make an
-additional hypercall that passes a new region of some size to DPM. After restarting the original
-hypercall, DPM will successfully find a region to hand to the EL2 service.
+\begin{figure}[h!tbp]
+\centering
+\captionsetup{justification=centering}
+\includegraphics{dpmmetadata}
+\caption{FlexAlloc: DPM Metadata Organization}
+\label{fig:dpmmetadata}
+\end{figure}
 
 If the host is under a lot of memory pressure, the host can request that DPM give control
-of available regions back to the host. The host passes a page that DPM will populate with
+of available regions back to the host. The host does so by invoking a hypercall that passes a page $P$, which DPM will populate with
 an array of metadata pertaining to regions that it is handing back. DPM will first unmap this
-page from the host. DPM will then search through the list for all available regions and add them
-to the page. DPM will finally remap the return page to the host and return. The host can then
-call \textit{free\_pages} on all of the return regions.
+page $P$ from the host. DPM will then search through the list mentioned above for all available regions and add them
+to $P$. Finally, DPM will remap $P$ to the host and return. The host can then
+call \textit{free\_pages} on all of the regions listed in $P$.
 
-In order for DPM to receive pages that it needs for it's own metadata, if the host receives an
-error when trying to hand a region of some size to DPM, then the host will know that it needs
-to pass a region of size \textit{PAGE\_SIZE} first to DPM in a seperate hypercall and then retry the
-original region pass hypercall before finally retrying the original hypercall pertaining
-to some other EL2 service. When DPM receives a region and notices that it is out of metadata
-space, DPM will search for a region of \textit{PAGE\_SIZE} to claim for itself before processing
+It is important to ensure that DPM can receive the pages that it needs for its own metadata.
+If the host receives an error when trying to hand a region of some size to DPM, the host will know that it needs
+to first pass a region of size \textit{PAGE\_SIZE} to DPM in a seperate hypercall, and then retry the
+original region pass hypercall, before finally retrying the original hypercall pertaining
+to some other EL2 service. 
+When DPM receives a region and notices that it is out of metadata
+space, it will search for a region of \textit{PAGE\_SIZE} to claim for itself before processing
 the larger region. If one is found, then DPM will claim it and process the region hand off
@@
 
-\subsection{Dynamic Page Manager - Considerations and Drawbacks}
-DPM allows for arbitrary request of regions from EL2 which is valuable when we may or may not
-need additional pages. If we know some hypercall will require EL2 to receive more pages,
+\subsubsection{Considerations and Drawbacks}
+DPM allows services within EL2 to request arbitrary regions of contiguous memory
+at arbitrary points in time. This is valuable when we may or may not need additional pages.
+If we know some hypercall will definitely require EL2 to receive more pages,
 the error-check-retry pathway may be an unnecessary cost. For example, when we want to
-boot a new VM, we know EL2 will need a stage 2 region for the VM. However, since DPM allows
-regions to be reused after they are made available again, a new VM launch can reuse the
-region from a previously shutdown VM. DPM is designed make it convenient to program against,
-but it’s generalized strategy can lead to extra, avoidable work. We designed DPM with
-programmability in mind over efficiency in hopes that it would make full dynamic allocation
-attainable for SeKVM services by avoiding excessively customized hand offs between the host
-and corevisor. Although we only partially implemented DPM, it would be interesting to
+boot a new VM, we know EL2 will definitely need a memory region for the stage-2 page table of the VM.
+However, since regions returned to DPM can be re-allocated to other EL2 services as long
+as memory is not reclaimed by the host, a new VM could potentially reuse the
+region returned by a previously-shutdown VM. DPM is designed to be convenient to use,
+but its generalized strategy can lead to extra, avoidable work. We design DPM with
+programmability in mind over efficiency, in hopes that it would make full dynamic allocation
+attainable for SeKVM services by avoiding excessively-customized handoffs between the host
+and corevisor. Although we only partially implement DPM, it would be interesting to
 compare the real-world performance metrics between our current stage-2 dynamic allocation
-strategy and a stage-2 dynamic allocation strategy built on top of DPM and a
+strategy, and a stage-2 dynamic allocation strategy built on top of DPM and a
 check-error-retry methodology.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%  LocalWords:  endnotes includegraphics fread ptr nobj noindent
-%%  LocalWords:  pdflatex acks
-
-%-------------------------------------------------------------------------------
-% \section{Floating Figures and Lists}
-% \label{sec:figs}
-%-------------------------------------------------------------------------------
-
-%---------------------------
-% \begin{figure}
-% \begin{center}
-% \begin{tikzpicture}
-%   \draw[thin,gray!40] (-2,-2) grid (2,2);
-%   \draw[<->] (-2,0)--(2,0) node[right]{$x$};
-%   \draw[<->] (0,-2)--(0,2) node[above]{$y$};
-%   \draw[line width=2pt,blue,-stealth](0,0)--(1,1)
-%         node[anchor=south west]{$\boldsymbol{u}$};
-%   \draw[line width=2pt,red,-stealth](0,0)--(-1,-1)
-%         node[anchor=north east]{$\boldsymbol{-u}$};
-% \end{tikzpicture}
-% \end{center}
-% \caption{\label{fig:vectors} Text size inside figure should be as big as
-%   caption's text. Text size inside figure should be as big as
-%   caption's text. Text size inside figure should be as big as
-%   caption's text. Text size inside figure should be as big as
-%   caption's text. Text size inside figure should be as big as
-%   caption's text. }
-% \end{figure}
-%% %---------------------------
-
-% Here's a typical reference to a floating figure:
-% Figure~\ref{fig:vectors}. Floats should usually be placed where latex
-% wants then. Figure\ref{fig:vectors} is centered, and has a caption
-% that instructs you to make sure that the size of the text within the
-% figures that you use is as big as (or bigger than) the size of the
-% text in the caption of the figures. Please do. Really.
-
-% In our case, we've explicitly drawn the figure inlined in latex, to
-% allow this tex file to cleanly compile. But usually, your figures will
-% reside in some file.pdf, and you'd include them in your document
-% with, say, \textbackslash{}includegraphics.
-
-% Lists are sometimes quite handy. If you want to itemize things, feel
-% free:
-
-% \begin{description}
-
-% \item[fread] a function that reads from a \texttt{stream} into the
-%   array \texttt{ptr} at most \texttt{nobj} objects of size
-%   \texttt{size}, returning returns the number of objects read.
-
-% \item[Fred] a person's name, e.g., there once was a dude named Fred
-%   who separated usenix.sty from this file to allow for easy
-%   inclusion.
-% \end{description}
